---
title: "R Notebook"
output:
  word_document: default
  html_notebook: default
---

```{r}
data <- read.csv("ptg_data.csv")
library(tidyverse)
library(dplyr)
library(ggplot2)
library(Hmisc)
library(knitr)
library(lavaan)
library(lcmm)
```

Descriptives and Correlations
```{r}
data1 <- dplyr::select(data, SUBJID, ss1.Age, ss1.Sex, ss1.Race, wave1_ptg, wave2_ptg, wave3_ptg, wave4_ptg, wave1_pts, wave2_pts, wave3_pts, wave1_etv, wave2_etv, wave3_etv)

data1$ss1.Race <- factor(data1$ss1.Race,
                         levels = c(1,2,3,4,5,6),
                         labels = c("Black or African American", "Hispanic or Latino/a", "Asian or Pacific Islander","White or Caucasian","Mixed Race","Other"))
data1 <- data1 %>% mutate(ss1.Sex=recode(ss1.Sex,'1'='1','2'='0'))
data1 <- data1 %>% mutate(ss1.Age=as.numeric(ss1.Age), na.rm = TRUE)
df_long <- data1 %>%
  pivot_longer(
    cols = -c(SUBJID, `ss1.Age`, `ss1.Sex`, `ss1.Race`), # Exclude SUBJID from reshaping
    names_to = c("wave", ".value"),  # Specify where parts of the old column names will go
    names_pattern = "(wave\\d+)_(ptg)"  # Use a regular expression to match old column names
  )
df_clean <- na.omit(df_long)
df_numeric <- df_clean
# Remove the letters from the 'SUBJID' column + wave column
df_numeric$SUBJID <- as.numeric(gsub("[A-Za-z]", "", df_numeric$SUBJID))
```
```{r}
view(df_long)
```
### MD: Intercept-only model
```{r}
imodel <- "Intercept =~ 1*walden_0 + 1*walden_1 + 1*walden_2 + 1*walden_3"
ifit <- growth(imodel, data=pathwayswide)
summary(ifit, fit.measures=T)

```

```{r}
m1a <- "Intercept =~ 1*wave1_ptg + 1*wave2_ptg + 1*wave3_ptg + 1*wave4_ptg"
m1a_fit <- growth(m1a, data=data1)
summary(m1a_fit, fit.measures=T)
```


##Model 1 = Intercept Only, variance controlled at 0
```{r}
m1b <- '
i =~ 1*wave1_ptg + 1*wave2_ptg + 1*wave3_ptg + 1*wave4_ptg
i ~~ 0*i
wave1_ptg ~~ r*wave1_ptg 
wave2_ptg ~~ r*wave2_ptg 
wave3_ptg ~~ r*wave3_ptg
wave4_ptg ~~ r*wave4_ptg
'
m1b_fit <- growth(m1b, data = data1)
summary(m1b_fit)
parameterestimates(m1b_fit, standardized = TRUE)
fitmeasures(m1b_fit)
```

Model 2 = Intercept only, variance > 0, random intercept model (models that individuals will have different intercepts)
```{r}
m2 <- '
i =~ 1*wave1_ptg + 1*wave2_ptg + 1*wave3_ptg + 1*wave4_ptg
wave1_ptg ~~ r*wave1_ptg 
wave2_ptg ~~ r*wave2_ptg 
wave3_ptg ~~ r*wave3_ptg
wave4_ptg ~~ r*wave4_ptg'
m2_fit <- growth(m2, data = data1)
summary(m2_fit)
parameterestimates(m2_fit, standardized = TRUE)
fitmeasures(m2_fit)
```

### MD Linear model
```{r}
smodel <- "Intercept =~ 1*walden_0 + 1*walden_1 + 1*walden_2 + 1*walden_3 
Slope =~ 0*walden_0 + .1*walden_1 + .2*walden_2 + .3*walden_3"
sfit <- growth(smodel, data=pathwayswide)
summary(sfit, fit.measures=T)
#this is a text comment within code.
```
```{r}
m3a <- "Intercept =~ 1*wave1_ptg + 1*wave2_ptg + 1*wave3_ptg + 1*wave4_ptg 
Slope =~ 0*wave1_ptg + .1*wave2_ptg + .2*wave3_ptg + .3*wave4_ptg"
m3a_fit <- growth(m3a, data=data1)
summary(m3a_fit, fit.measures=T)
```

Model 3 = Random slope
```{r}
m3b <- '
i =~ 1*wave1_ptg + 1*wave2_ptg + 1*wave3_ptg + 1*wave4_ptg
s =~ 0*wave1_ptg + 1*wave2_ptg + 2*wave3_ptg + 3*wave4_ptg
s ~ 0*1
s ~~ 0*i
wave1_ptg ~~ r*wave1_ptg 
wave2_ptg ~~ r*wave2_ptg 
wave3_ptg ~~ r*wave3_ptg
wave4_ptg ~~ r*wave4_ptg

m3b_fit <- growth(m3b, data = data1)
summary(m3b_fit)
parameterestimates(m3b_fit, standardized = TRUE)
fitmeasures(m3b_fit)
```
Model 4 = Allow slope to vary
```{r}
m4 <- '
i =~ 1*wave1_ptg + 1*wave2_ptg + 1*wave3_ptg + 1*wave4_ptg
s =~ 0*wave1_ptg + 1*wave2_ptg + 2*wave3_ptg + 3*wave4_ptg
wave1_ptg ~~ r*wave1_ptg 
wave2_ptg ~~ r*wave2_ptg 
wave3_ptg ~~ r*wave3_ptg
wave4_ptg ~~ r*wave4_ptg
'
m4_fit <- growth(m4, data = data1)
summary(m4_fit)
parameterestimates(m4_fit, standardized = TRUE)
fitmeasures(m4_fit)
```
Model 5 = Allow residuals to vary 
```{r}
m5 <- '
i =~ 1*wave1_ptg + 1*wave2_ptg + 1*wave3_ptg + 1*wave4_ptg
s =~ 0*wave1_ptg + 1*wave2_ptg + 2*wave3_ptg + 3*wave4_ptg
'
m5_fit <- growth(m5, data = data1)
summary(m5_fit)
parameterestimates(m5_fit, standardized = TRUE)
fitmeasures(m5_fit)
```

### MD Quadratic model
This is not code, so we can type normally without a hashtag.
```{r}
qmodel <- "Intercept =~ 1*walden_0 + 1*walden_1 + 1*walden_2 + 1*walden_3 
Slope =~ 0*walden_0 + .1*walden_1 + .2*walden_2 + .3*walden_3 
Quadratic =~ 0*walden_0 + .01*walden_1 + .04*walden_2 + .09*walden_3"
qfit <- growth(qmodel, data=pathwayswide)
summary(qfit, fit.measures=T)
```

Model 6 = quadratic growth model
```{r}
m6 <- "Intercept =~ 1*wave1_ptg + 1*wave2_ptg + 1*wave3_ptg + 1*wave4_ptg 
Slope =~ 0*wave1_ptg + 1*wave2_ptg + 2*wave3_ptg + 3*wave4_ptg
Quadratic =~ 0*wave1_ptg + .01*wave2_ptg + .04*wave3_ptg + .09*wave4_ptg"
m6_fit <- growth(m6, data=data1)
summary(m6_fit, fit.measures=T)
```


```{r}
ggplot(df_clean, aes(x=wave, y=ptg, group = SUBJID)) +
  geom_point(size = .1) +
  geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess")
```
```{r}
ggplot(df_numeric, aes(x=wave, y=ptg, group=SUBJID, colour=SUBJID)) +
  geom_point(size=.1) +
  geom_line(size=.1) +
  geom_smooth(se=FALSE, method="loess", aes(colour=NULL))

```





```{r}
class(data1)
dim(data1) 
names(data1) 
str(data1) 
glimpse(data1) 
summary(data1)
```


```{r}
data1$ss1.Race
```
```{r}
ggplot(data1, aes(x = ss1.Race)) +
  geom_bar(fill = "pink", color = "black") +
  labs(x = "Race", y = "Count") +
  theme_minimal()
```
```{r}
df_long <- data1 %>%
  pivot_longer(
    cols = -SUBJID,  # Exclude SUBJID from reshaping
    names_to = c("wave", ".value"),  # Specify where parts of the old column names will go
    names_pattern = "(wave\\d+)_(pts|ptg)"  # Use a regular expression to match old column names
  )

```
```{r}
df_long
```


```{r}
df_long_cleaned <- na.omit(df_long)
```
```{r}
df_long_cleaned
```



```{r}
library(dplyr)
library(ggplot2)

# Calculate the mean ptg at each wave
mean_ptg_by_wave <- df_long %>%
  group_by(wave) %>%
  summarise(mean_ptg = mean(ptg, na.rm = TRUE))
```
```{r}
mean_ptg_by_wave
```

```{r}
ggplot(mean_ptg_by_wave, aes(x = wave, y = mean_ptg)) +
  geom_col()
```
```{r}
data3 <- na.omit(data.frame(data1$wave1_ptg, data1$wave2_ptg, data1$wave3_ptg))
```
```{r}
ptg.cov <- cov(data3)
```
```{r}
data3
```



```{r}
mean_ptg_by_wave
```
```{r}
ptg_mean <- c(3.35, 3.45, 3.48)
```




```{r}
ggplot(data,aes(x=wave1_ptg)) +   geom_histogram(binwidth=.4) 
```
```{r}
ggplot(data,aes(x=wave2_ptg)) +   geom_histogram(binwidth = .4) 
```
```{r}
ggplot(data,aes(x=wave3_ptg)) +   geom_histogram(binwidth = .4) 
```
```{r}
ggplot(data,aes(x=wave1_pts)) +   geom_histogram(binwidth = .2) 
```
```{r}
ggplot(data,aes(x=wave1_etv)) +   geom_histogram(binwidth = .2) 
```

```{r}
df <- data1 %>% select(c(5:13))
data2 <- data1 %>% 
  mutate(across(contains("wave"), scale))
```
```{r}
data2
```
```{r}
corstars <-function(x, method=c("pearson", "spearman"), removeTriangle=c("upper", "lower"),
                     result=c("none", "html", "latex")){
    #Compute correlation matrix
    require(Hmisc)
    x <- as.matrix(x)
    correlation_matrix<-rcorr(x, type=method[1])
    R <- correlation_matrix$r # Matrix of correlation coeficients
    p <- correlation_matrix$P # Matrix of p-value 
    
    ## Define notions for significance levels; spacing is important.
    mystars <- ifelse(p < .0001, "****", ifelse(p < .001, "*** ", ifelse(p < .01, "**  ", ifelse(p < .05, "*   ", "    "))))
    
    ## trunctuate the correlation matrix to two decimal
    R <- format(round(cbind(rep(-1.11, ncol(x)), R), 2))[,-1]
    
    ## build a new matrix that includes the correlations with their apropriate stars
    Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x))
    diag(Rnew) <- paste(diag(R), " ", sep="")
    rownames(Rnew) <- colnames(x)
    colnames(Rnew) <- paste(colnames(x), "", sep="")
    
    ## remove upper triangle of correlation matrix
    if(removeTriangle[1]=="upper"){
      Rnew <- as.matrix(Rnew)
      Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
      Rnew <- as.data.frame(Rnew)
    }
    
    ## remove lower triangle of correlation matrix
    else if(removeTriangle[1]=="lower"){
      Rnew <- as.matrix(Rnew)
      Rnew[lower.tri(Rnew, diag = TRUE)] <- ""
      Rnew <- as.data.frame(Rnew)
    }
    
    ## remove last column and return the correlation matrix
    Rnew <- cbind(Rnew[1:length(Rnew)-1])
    if (result[1]=="none") return(Rnew)
    else{
      if(result[1]=="html") print(xtable(Rnew), type="html")
      else print(xtable(Rnew), type="latex") 
    }
} 
```
```{r}
corstars(df)
```
```{r}
pc1_ptg <- lm(wave2_ptg ~ wave1_ptg + wave1_pts, data = data1)
pc2_ptg <- lm(wave3_ptg ~ wave2_ptg + wave2_pts, data = data1)
pc1_pts <- lm(wave2_pts ~ wave1_ptg + wave1_pts, data = data1)
pc2_pts <- lm(wave3_pts ~ wave2_ptg + wave2_pts, data = data1)
```
```{r}
summary(pc1_pts)
```


```{r}
cortable <- corstars(df)
```

```{r}
class(cortable)
```
```{r}
write.table(cortable, file = '/Users/kaylisealgrim/Documents/cors.txt', sep = ",")
```




```{r}
# specify model
model <- '
    # latent variables
    i =~ 1*wave1_ptg + 1*wave2_ptg + 1*wave3_ptg
    s =~ 0*wave1_ptg + 1*wave2_ptg + 2*wave3_ptg

    # regression on latent variables
    wave3_ptg ~ i + s

    # specify means and variances of latent variables (intercepts and slopes)
    i ~~ 1*i
    s ~~ 0*s
    i ~~ 0*s
'

# fit model
fit <- sem(model, data = data1)

# get summary of model fit
summary(fit, fit.measures=TRUE)

```


1. Latent growth curve

```{r}
# Specify the LGC model
model <- '
    # Latent variables
    i =~ 1*wave1_ptg + 1*wave2_ptg + 1*wave3_ptg + 1*wave4_ptg
    s =~ 0*wave1_ptg + 1*wave2_ptg + 2*wave3_ptg + 3*wave4_ptg
'

# Fit the LGC model
fit <- growth(model, data = data1)

# Summarize the results
summary(fit, fit.measures =T)
```
2. Quadratix 

q=~ *0, *1, *4, *9


```{r}
# Define the model for ETV
qmodel <- '
  # Latent intercept and slope
  i =~ 1*wave1_ptg + 1*wave2_ptg + 1*wave3_ptg + 1*wave4_ptg 
  s =~ 0*wave1_ptg + 1*wave2_ptg + 2*wave3_ptg + 3*wave4_ptg 
  q =~ 0*wave1_ptg + 1*wave2_ptg + 4*wave3_ptg + 9*wave4_ptg 
'
q_fit <- growth(qmodel, data = data1)
summary(q_fit, fit.measures=T)
```

In this model: i represents the initial level (intercept) of PTG, and s represents the rate of change (slope) in PTG over time. The lines i =~ 1*wave1_ptg + 1*wave2_ptg + 1*wave3_ptg and s =~ 0*wave1_ptg + 1*wave2_ptg + 2*wave3_ptg specify the measurement part of the model. For the slope, the coefficients before each wave represent the time scores and determine the shape of the growth trajectory. In this case, a linear growth trajectory is specified. The lines i ~~ 1*i and s ~~ 1*s estimate the variances of the intercept. 
```{r}
# Define the model for PTS
model_pts <- '
  # Latent intercept and slope
  i =~ 1*wave1_ptg + 1*wave2_ptg + 1*wave3_ptg
  s =~ 0*wave1_ptg + 1*wave2_ptg + 2*wave3_ptg
  
  # Regressions
  i ~ wave1_pts
  s ~ wave1_pts
  
  # Variances
  i ~~ i
  s ~~ s
'

fit_pts <- sem(model_pts, data = data2)
summary(fit_pts)
```

```{r}
summary(data1)
```








```{r}
df_long <- data1 %>%
  pivot_longer(
    cols = -SUBJID,  # Exclude SUBJID from reshaping
    names_to = c("wave", ".value"),  # Specify where parts of the old column names will go
    names_pattern = "(wave\\d+)_(pts|ptg)"  # Use a regular expression to match old column names
  )

```
```{r}
df_long
```
```{r}
class(dat)
dim(dat) 
names(dat) 
str(dat) 
glimpse(dat) 
summary(dat) 
```
i. Example ri-clpm from Mulder & Hamaker
```{r}
dat <-read.table("RICLPM.dat",
                 col.names = c(
                   "x1", "x2", "x3", "x4", "x5",
                   "y1", "y2", "y3", "y4", "y5"
                 ))
```

```{r}
RICLPM <- '
# Create between components (random intercepts)
  RIx =~ 1*x1 + 1*x2 + 1*x3 + 1*x4 + 1*x5
  RIy =~ 1*y1 + 1*y2 + 1*y3 + 1*y4 + 1*y5
  
  # Create within-person centered variables
  wx1 =~ 1*x1
  wx2 =~ 1*x2
  wx3 =~ 1*x3 
  wx4 =~ 1*x4
  wx5 =~ 1*x5
  wy1 =~ 1*y1
  wy2 =~ 1*y2
  wy3 =~ 1*y3
  wy4 =~ 1*y4
  wy5 =~ 1*y5
  
    # Estimate lagged effects between within-person centered variables
  wx2 + wy2 ~ wx1 + wy1
  wx3 + wy3 ~ wx2 + wy2
  wx4 + wy4 ~ wx3 + wy3
  wx5 + wy5 ~ wx4 + wy4
  
   # Estimate covariance between within-person centered variables at first wave
  wx1 ~~ wy1 # Covariance
  
  # Estimate covariances between residuals of within-person centered variables 
  # (i.e., innovations)
  wx2 ~~ wy2
  wx3 ~~ wy3
  wx4 ~~ wy4
  wx5 ~~ wy5
  
  # Estimate variance and covariance of random intercepts
  RIx ~~ RIx
  RIy ~~ RIy
  RIx ~~ RIy

  # Estimate (residual) variance of within-person centered variables
  wx1 ~~ wx1 # Variances
  wy1 ~~ wy1 
  wx2 ~~ wx2 # Residual variances
  wy2 ~~ wy2 
  wx3 ~~ wx3 
  wy3 ~~ wy3 
  wx4 ~~ wx4 
  wy4 ~~ wy4 
  wx5 ~~ wx5
  wy5 ~~ wy5'
RICLPM.fit <- lavaan(RICLPM,
  data = dat, 
  missing = "ML", 
  meanstructure = T, 
  int.ov.free = T
)
summary(RICLPM.fit, standardized = T)
``'

```{r}
data2 <- data1[c('wave1_pts', 'wave2_pts', 'wave3_pts')]
icc_result1 <- icc(data2, model = "oneway", type = "agreement", unit = "single")
print(icc_result1)
```
```{r}
data3 <- data1[c('wave1_ptg', 'wave2_ptg', 'wave3_ptg')]
icc_result2 <- icc(data3, model = "oneway", type = "agreement", unit = "single")
print(icc_result)
```


1. Random intercept cross-lagged panel model

```{r}
model1 <- '
  # between components (random intercepts)
  RIpts =~ 1*wave1_pts + 1*wave2_pts + 1*wave3_pts
  RIptg =~ 1*wave1_ptg + 1*wave2_ptg + 1*wave3_ptg
  
  # within person centered variables
  wpts1 =~ 1*wave1_pts
  wpts2 =~ 1*wave2_pts
  wpts3 =~ 1*wave3_pts
  wptg1 =~ 1*wave1_ptg
  wptg2 =~ 1*wave2_ptg
  wptg3 =~ 1*wave3_ptg

   # Estimate lagged effects between within-person centered variables
   wpts2 + wptg2 ~ wpts1 + wptg1
   wpts3 + wptg3 ~ wpts2 + wptg2
  
  # Estimate covariance between within-person centered variables at first wave
  wpts1 ~~ wptg1
  
    # Estimate covariances between residuals of within-person centered variables 
  # (i.e., innovations)
 wpts2 ~~ wptg2
  wpts3 ~~ wptg3
  
  
  #cross-lagged paths
  wave2_pts ~ c21*wave1_ptg
  wave3_pts ~ c21*wave2_ptg 
  wave2_ptg ~ c12*wave1_pts
  wave3_ptg ~ c12*wave2_pts
  
  # Estimate variance and covariance of random intercepts
  RIpts ~~ RIptg

 # Estimate (residual) variance of within-person centered variables
  wpts1 ~~ wpts1 # Variances
  wptg1 ~~ wptg1 
  wpts2 ~~ wpts2 # Residual variances
  wptg2 ~~ wptg2 
  wpts3 ~~ wpts3 
  wptg3 ~~ wptg3'

ptg.fit <- lavaan(model1,
  data = data1, 
  missing = "ML", 
  meanstructure = T, 
  int.ov.free = T
)
summary(ptg.fit, standardized = T)

```

2. Latent growth curve
```{r}
data1
```
```{r}
# Specify the model syntax
model <- '
  # Latent growth factors
  PTG =~ wave1_pts + wave2_pts + wave3_pts
  PTS =~ wave1_ptg + wave2_ptg + wave3_ptg
  ETV =~ wave1_etv + wave2_pts + wave3_pts

  # Regression paths
  PTG ~ PTS
  PTS ~ ETV

  # Residual variances
  wave1_ptg ~~ wave1_ptg
  wave2_ptg ~~ wave2_ptg
  wave3_ptg ~~ wave3_ptg
  wave1_pts ~~ wave1_pts
  wave2_pts ~~ wave2_pts
  wave3_pts ~~ wave3_pts
  wave1_etv ~~ wave1_etv
  wave2_etv ~~ wave2_etv
  wave3_etv ~~ wave3_etv
'

# Fit the model to the data
fit <- lavaan(model, data = data1, missing = "FIML")

# Print the model results
summary(fit)
```

```{r}
df_long <- data2 %>%
  pivot_longer(
    cols = -SUBJID,  # Exclude SUBJID from reshaping
    names_to = c("wave", ".value"),  # Specify where parts of the old column names will go
    names_pattern = "(wave\\d+)_(.*)"  # Use a regular expression to match old column names
  )

```
```{r}
df_long
```




3. Latent class mixture model

```{r}
data1
```
```{r}
df_long <- data1 %>%
  pivot_longer(
    cols = -SUBJID,  # Exclude SUBJID from reshaping
    names_to = c("wave", ".value"),  # Specify where parts of the old column names will go
    names_pattern = "(wave\\d+)_(pts|ptg)"  # Use a regular expression to match old column names
  )

```
```{r}
df_numeric <- df_long
# Remove the letters from the 'SUBJID' column + wave column
df_numeric$SUBJID <- as.numeric(gsub("[A-Za-z]", "", df_numeric$SUBJID))
df_numeric$wave <- as.numeric(gsub("[A-Za-z]", "", df_numeric$wave))
df_numeric
```

```{r}
class(data1)
dim(df_numeric) 
names(df_numeric) 
str(data1) 
glimpse(df_numeric) 
summary(df_numeric) 
```
----
```{r}
mpts1 <- lcmm(pts ~ wave, random = ~ wave, subject = 'SUBJID', ng = 1, data = df_numeric)
mpts2 <- lcmm(pts ~ wave, random = ~ wave, subject = 'SUBJID', mixture = ~ wave, ng = 2, data = df_numeric)
mpts3 <- lcmm(pts ~ wave, random = ~ wave, subject = 'SUBJID', mixture = ~ wave, ng = 3, data = df_numeric)
mpts4 <- lcmm(pts ~ wave, random = ~ wave, subject = 'SUBJID', mixture = ~ wave, ng = 4, data = df_numeric)
mpts5 <- lcmm(pts ~ wave, random = ~ wave, subject = 'SUBJID', mixture = ~ wave, ng = 5, data = df_numeric)
```
```{r}
mptg1 <- lcmm(ptg ~ wave, random = ~ wave, subject = 'SUBJID', ng = 1, data = df_numeric)
mptg2 <- lcmm(ptg ~ wave, random = ~ wave, subject = 'SUBJID', mixture = ~ wave, ng = 2, data = df_numeric)
mptg3 <- lcmm(ptg ~ wave, random = ~ wave, subject = 'SUBJID', mixture = ~ wave, ng = 3, data = df_numeric)
mptg4 <- lcmm(ptg ~ wave, random = ~ wave, subject = 'SUBJID', mixture = ~ wave, ng = 4, data = df_numeric)
mptg5 <- lcmm(ptg ~ wave, random = ~ wave, subject = 'SUBJID', mixture = ~ wave, ng = 5, data = df_numeric)
```


```{r}
#comparing model fit for pts trajectories
summary(mpts1)
summary(mpts2)
summary(mpts3)
summary(mpts4)
summary(mpts5)
```
```{r}
#comparing model fit for ptg trajectories
summary(mptg1)
summary(mptg2)
summary(mptg3)
summary(mptg4)
summary(mptg5)
```

Results: 

There were slight differences in AIC and BIC when comparing post-traumatic stress trajectories. In post-traumatic growth, a two-class model showed the best fit for both AIC and BIC.

```{r}
model <- gbmt(x.names = c('pts', 'ptg'), unit = 'wave', time = 'wave', ng = 5, data = df_long)
```


```{r}
ggplot(data = df_numeric, aes(x = pts, y = ptg, color = factor(wave))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "PTS", y = "PTG", color = "Wave") +
  theme_minimal()

```
```{r}
df <- data1 %>% select(c(2:7))
```
```{r}
corstars(df)
```
```{r}
corstars <-function(x, method=c("pearson", "spearman"), removeTriangle=c("upper", "lower"),
                     result=c("none", "html", "latex")){
    #Compute correlation matrix
    require(Hmisc)
    x <- as.matrix(x)
    correlation_matrix<-rcorr(x, type=method[1])
    R <- correlation_matrix$r # Matrix of correlation coeficients
    p <- correlation_matrix$P # Matrix of p-value 
    
    ## Define notions for significance levels; spacing is important.
    mystars <- ifelse(p < .0001, "****", ifelse(p < .001, "*** ", ifelse(p < .01, "**  ", ifelse(p < .05, "*   ", "    "))))
    
    ## trunctuate the correlation matrix to two decimal
    R <- format(round(cbind(rep(-1.11, ncol(x)), R), 2))[,-1]
    
    ## build a new matrix that includes the correlations with their apropriate stars
    Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x))
    diag(Rnew) <- paste(diag(R), " ", sep="")
    rownames(Rnew) <- colnames(x)
    colnames(Rnew) <- paste(colnames(x), "", sep="")
    
    ## remove upper triangle of correlation matrix
    if(removeTriangle[1]=="upper"){
      Rnew <- as.matrix(Rnew)
      Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
      Rnew <- as.data.frame(Rnew)
    }
    
    ## remove lower triangle of correlation matrix
    else if(removeTriangle[1]=="lower"){
      Rnew <- as.matrix(Rnew)
      Rnew[lower.tri(Rnew, diag = TRUE)] <- ""
      Rnew <- as.data.frame(Rnew)
    }
    
    ## remove last column and return the correlation matrix
    Rnew <- cbind(Rnew[1:length(Rnew)-1])
    if (result[1]=="none") return(Rnew)
    else{
      if(result[1]=="html") print(xtable(Rnew), type="html")
      else print(xtable(Rnew), type="latex") 
    }
} 
```
```{r}
pc1 <- lm(wave2_pts ~ wave1_pts + wave1_ptg, data = data1)
pc2 <- lm(wave2_ptg ~ wave1_pts + wave1_ptg, data = data1)
pc3 <- lm(wave3_pts ~ wave2_pts + wave2_ptg, data = data1)
pc4 <- lm(wave3_ptg ~ wave2_pts + wave2_ptg, data = data1)
```
```{r}
summary(pc1)
summary(pc2)
summary(pc3)
summary(pc4)
```








