---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
  word_document: default
---
data - complete dataset
data1 - variables of interest for all waves
df_var - all numeric variables for correlations
df_long - variables of interest in long format
df_clean - long format dataset with NAs removed 
df_pred - pts and etv at time 1
df_numeric - df_clean with SUBJID and wave in numeric form to analyze


```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(knitr)
library(lavaan)
library(lcmm)
library(lme4)
library(nnet)
library(chisq.posthoc.test)
corstars <-function(x, method=c("pearson", "spearman"), removeTriangle=c("upper", "lower"),
                     result=c("none", "html", "latex")){
    #Compute correlation matrix
    require(Hmisc)
    x <- as.matrix(x)
    correlation_matrix<-rcorr(x, type=method[1])
    R <- correlation_matrix$r # Matrix of correlation coeficients
    p <- correlation_matrix$P # Matrix of p-value 
    
    ## Define notions for significance levels; spacing is important.
    mystars <- ifelse(p < .0001, "****", ifelse(p < .001, "*** ", ifelse(p < .01, "**  ", ifelse(p < .05, "*   ", "    "))))
    
    ## trunctuate the correlation matrix to two decimal
    R <- format(round(cbind(rep(-1.11, ncol(x)), R), 2))[,-1]
    
    ## build a new matrix that includes the correlations with their apropriate stars
    Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x))
    diag(Rnew) <- paste(diag(R), " ", sep="")
    rownames(Rnew) <- colnames(x)
    colnames(Rnew) <- paste(colnames(x), "", sep="")
    
    ## remove upper triangle of correlation matrix
    if(removeTriangle[1]=="upper"){
      Rnew <- as.matrix(Rnew)
      Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
      Rnew <- as.data.frame(Rnew)
    }
    
    ## remove lower triangle of correlation matrix
    else if(removeTriangle[1]=="lower"){
      Rnew <- as.matrix(Rnew)
      Rnew[lower.tri(Rnew, diag = TRUE)] <- ""
      Rnew <- as.data.frame(Rnew)
    }
    
    ## remove last column and return the correlation matrix
    Rnew <- cbind(Rnew[1:length(Rnew)-1])
    if (result[1]=="none") return(Rnew)
    else{
      if(result[1]=="html") print(xtable(Rnew), type="html")
      else print(xtable(Rnew), type="latex") 
    }
} 
```


Naming
```{r}
data <- read.csv("ptg_data.csv")
data <- data %>% 
  mutate(wave4_pts = rowMeans(select(.,ss4.pts01,ss4.pts02,ss4.pts03,ss4.pts04,ss4.pts05,ss4.pts06,ss4.pts07,ss4.pts08,ss4.pts09)))
data <- data %>% 
  mutate(wave4_ptg = rowMeans(select(.,ss4.ptg01,ss4.ptg02,ss4.ptg03,ss4.ptg04,ss4.ptg05,ss4.ptg06,ss4.ptg07,ss4.ptg08,ss4.ptg09,ss4.ptg10)))
data <- data %>% 
  mutate(wave4_etv = rowMeans(select(.,ss4.NVL01,ss4.NVL02,ss4.NVL03,ss4.NVL04,ss4.NVL05,ss4.NVL06,ss4.NVL07,ss4.NVL08,ss4.NVL09,ss4.NVL10,ss4.NVL11,
                                     ss4.NVL12,ss4.NVL13,ss4.NVL14,ss4.NVL15,ss4.NVL16,ss4.NVL17,ss4.NVL18,ss4.NVL19,ss4.NVL20)))
data <- data %>% 
  mutate(wave1_gun = rowMeans(select(.,ss1.NVL01,ss1.NVL03,ss1.NVL07,ss1.NVL08,ss1.NVL14,ss1.NVL15)))
data <- data %>% 
  mutate(wave2_gun = rowMeans(select(.,ss2.NVL01,ss2.NVL03,ss2.NVL07,ss2.NVL08,ss2.NVL14,ss2.NVL15)))
data <- data %>% 
  mutate(wave3_gun = rowMeans(select(.,ss3.NVL01,ss3.NVL03,ss3.NVL07,ss3.NVL08,ss3.NVL14,ss3.NVL15)))
data1 <- data %>% select(c('SUBJID','ss1.Age','ss1.Sex','ss1.Race','ss2.Age','ss2.Sex','ss2.Race','ss3.Age','ss3.Sex','ss3.Race','wave1_ptg','wave2_ptg','wave3_ptg','wave1_pts','wave2_pts','wave3_pts','wave1_etv','wave2_etv','wave3_etv','wave4_pts','wave4_etv','wave4_ptg','ss4.Age','ss4.Sex','ss4.Race'))
##getting natural log for relevant variables


data1 <- data1 %>%
  mutate(across(matches("^wave\\d+_ptg$"), list(logptg = ~log(.))))

data1 <- data1 %>%
rename(wave1_logptg = wave1_ptg_logptg)
data1 <- data1 %>%
rename(wave2_logptg = wave2_ptg_logptg)
data1 <- data1 %>%
rename(wave3_logptg = wave3_ptg_logptg)
data1 <- data1 %>%
rename(wave4_logptg = wave4_ptg_logptg)

data2 <- data %>% select(c('SUBJID','ss1.Age','ss1.Sex','ss1.Race','wave1_ptg','wave2_ptg','wave3_ptg','wave1_pts','wave1_etv','wave1_gun'))
data1$ss1.Race <- factor(data1$ss1.Race,
                         levels = c(1,2,3,4,5,6),
                         labels = c("Black", "Hispanic", "Asian","White","Mixed","Other"))
data1$ss2.Race <- factor(data1$ss2.Race,
                         levels = c(1,2,3,4,5,6),
                         labels = c("Black", "Hispanic", "Asian","White","Mixed","Other"))
data1$ss3.Race <- factor(data1$ss3.Race,
                         levels = c(1,2,3,4,5,6),
                         labels = c("Black", "Hispanic", "Asian","White","Mixed","Other"))
data1$ss4.Race <- factor(data1$ss4.Race,
                         levels = c(1,2,3,4,5,6),
                         labels = c("Black", "Hispanic", "Asian","White","Mixed","Other"))
data1 <- data1 %>%
  mutate(race_attrition1 = if_else(ss1.Race == "White", 1, 0))
data1 <- data1 %>%
  mutate(race_attrition2 = if_else(ss2.Race == "White", 1, 0))
data1 <- data1 %>%
  mutate(race_attrition3 = if_else(ss3.Race == "White", 1, 0))
data1 <- data1 %>%
  mutate(race_attrition4 = if_else(ss4.Race == "White", 1, 0))



data1 <- data1 %>% mutate(ss1.Sex=recode(ss1.Sex,'1'='1','2'='0'))
data1 <- data1 %>% mutate(ss1.Age=as.numeric(ss1.Age), na.rm = TRUE)
data1 <- data1 %>% mutate(ss2.Age=as.numeric(ss2.Age), na.rm = TRUE)
data1 <- data1 %>% mutate(ss3.Age=as.numeric(ss3.Age), na.rm = TRUE)
data1 <- data1 %>% mutate(Gender=as.numeric(ss1.Sex), na.rm = TRUE)
data1 <- data1 %>%
  mutate(category = ifelse(is.na(ss4.Sex), "No Data", "Has Data"))
# Relevel the factor to make '4' the reference category
data1$ss1.Race <- relevel(factor(data1$ss1.Race), ref = 'White')
mat <- model.matrix(~ ss1.Race - 1, data = data1)
df_dummies <- as.data.frame(mat)
data1 <- cbind(data1, df_dummies)
df_long <- data1 %>%
  pivot_longer(
    cols = -c(SUBJID, `ss1.Age`, `Gender`, `ss1.Race`,`category`,`ss1.RaceWhite`,`ss1.RaceBlack`,`ss1.RaceHispanic`,
              `ss1.RaceAsian`,`ss1.RaceMixed`,`ss1.RaceOther`), # Exclude SUBJID from reshaping
    names_to = c("wave", ".value"),  # Specify where parts of the old column names will go
    names_pattern = "(wave\\d+)_(ptg|pts|etv|logptg)"  # Use a regular expression to match old column names
  )
df_clean <- na.omit(df_long)
df_numeric <- df_clean
# Remove the letters from the 'SUBJID' column + wave column
df_numeric$SUBJID <- as.numeric(gsub("[A-Za-z]", "", df_numeric$SUBJID))
df_var <- data1 %>% 
  select("wave1_ptg","wave2_ptg","wave3_ptg","wave4_ptg","wave1_pts","wave2_pts","wave3_pts","wave4_pts","wave1_etv","wave2_etv","wave3_etv","wave4_etv")
df_numeric <- df_numeric %>%
  group_by(SUBJID) %>%
  mutate(etv1 = first(etv), pts1 = first(pts), ptg1 = first(ptg)) %>%
  ungroup()
df_numeric <- df_numeric %>%
  group_by(SUBJID) %>%
  mutate(pts3 = last(pts)) %>%
  ungroup()
df_numeric$wave <- as.numeric(sub("wave", "", df_numeric$wave))
lcga1_linear = hlme(fixed = ptg ~ wave,
                    subject = "SUBJID",
                    ng = 1,
                    data = df_numeric)
lcga3_linear = gridsearch(rep=10,
                          maxiter=10,
                          minit=lcga1_linear,  
                          hlme(ptg~wave,  
                               subject="SUBJID",
                               ng=3,
                               data=df_numeric,
                               mixture=~wave)) 
df_numeric <- merge(df_numeric, round(lcga3_linear$pprob, 3)[, 1:2], by = "SUBJID")
# Convert the 'class' column to a factor and store in 'clust3'
df_numeric$clust3 <- as.factor(df_numeric$class)

# Remove the 'class' column
df_numeric$class <- NULL
missing <- df_numeric %>% filter(wave == "1") %>% 
  select(category,pts1, ptg1,etv1,Gender,ss1.Race,ss1.Age,clust3) 
# Set "low" as the reference category for 'clust3'
#ADD this back in to the naming category
# Rename the factor levels of 'clust3'

levels_label <- df_numeric %>% 
  group_by(clust3) %>% 
  summarize(mean_ptg = mean(ptg, na.rm = TRUE))

levels_label <- levels_label %>% 
  arrange(desc(mean_ptg)) %>% 
  mutate(rank = row_number())

levels_label <- levels_label %>%
  mutate(label = case_when(
    rank == 1 ~ "high",
    rank == 2 ~ "mid",
    rank == 3 ~ "low"
  ))
df_numeric <- df_numeric %>%
  left_join(levels_label, by = "clust3") %>%
  mutate(clust3_label = label) %>%
  select(-mean_ptg, -rank, -label)
df_relabel <- data1
df_wide <- df_numeric %>%
  pivot_wider(names_from = wave, values_from = c(clust3, clust3_label), id_cols = SUBJID)
df_wide$SUBJID <- as.factor(df_wide$SUBJID)
df_relabel$SUBJID <- as.factor(df_relabel$SUBJID)
df_relabel$SUBJID <- as.numeric(gsub("[A-Za-z]", "", df_relabel$SUBJID))
merged_df <- merge(df_wide, df_relabel, by = "SUBJID")
select_df <- merged_df %>% select(SUBJID, clust3_1, clust3_label_1, wave1_etv, wave1_pts, Gender, ss1.RaceBlack, ss1.RaceHispanic, ss1.RaceAsian, ss1.RaceMixed, ss1.RaceOther, wave1_ptg, ss1.Race)
select_df <- select_df %>% rename(clust3 = clust3_1, clust3_label = clust3_label_1)
merge_model <- multinom(clust3_label ~ wave1_etv + wave1_pts + Gender + ss1.RaceBlack + ss1.RaceHispanic + ss1.RaceAsian + ss1.RaceMixed + ss1.RaceOther, data = select_df)
summary(merge_model)
df_wide_2 <- df_numeric %>%
  pivot_wider(names_from = wave, 
              values_from = c(clust3_label, etv1, pts1, Gender, ss1.RaceBlack, ss1.RaceHispanic, ss1.RaceAsian, ss1.RaceMixed, ss1.RaceOther, ptg),
              id_cols = SUBJID)
              
df_wide_2 <- df_wide_2 %>% 
  rename(clust3_label = clust3_label_1, etv1 = etv1_1, pts1 = pts1_1, Gender = Gender_1, ss1.RaceBlack = ss1.RaceBlack_1, ss1.RaceHispanic = ss1.RaceHispanic_1, ss1.RaceAsian = ss1.RaceAsian_1, ss1.RaceMixed = ss1.RaceMixed_1, ss1.RaceOther = ss1.RaceOther_1, ptg = ptg_1)

df_wide_2 <- df_wide_2 %>% 
  select(clust3_label, etv1, pts1, Gender, ss1.RaceBlack, ss1.RaceHispanic, ss1.RaceAsian, ss1.RaceMixed, ss1.RaceOther, ptg)
df_wide_2$clust3_label <- factor(df_wide_2$clust3_label, levels = c("low", "mid", "high"))
df_numeric$clust3_label <- factor(df_numeric$clust3_label, levels = c("low", "mid", "high"))
```


```{r}
library(haven)

write_sav(select_df, "select_data.sav")
```


```{r}
table(select_df$clust3_label)
```

```{r}
##The emergency block
view(select_df)
merge_model <- multinom(clust3_label ~ wave1_etv + wave1_pts + Gender + ss1.RaceBlack + ss1.RaceHispanic + ss1.RaceAsian + ss1.RaceMixed + ss1.RaceOther, data = select_df)
summary(merge_model)

#anovas a = cluster label, b = cluster class number; same results means I'm comforable using cluster labels
aov3a <- aov(wave1_etv ~ clust3_label, data = select_df)
summary(aov3a)
aov3b <- aov(wave1_etv ~ clust3, data = select_df)
summary(aov3b)
aov4a <- aov(wave1_pts ~ clust3_label, data = select_df)
summary(aov4a)
aov4b <- aov(wave1_pts ~ clust3, data = select_df)
summary(aov4b)
TukeyHSD(aov3a)
TukeyHSD(aov4a)
TukeyHSD(aov3b)
TukeyHSD(aov4b)
table_gender <- table(select_df$Gender, select_df$clust3_label)
table_race <- table(select_df$ss1.Race, select_df$clust3_label)
chisq.test(table_gender)
chisq.test(table_race)
table(select_df$ss1.race, select_df$clust3_label)
table(select_df$Gender, select_df$clust3_label) 
length(select_df$ss1.Race)

complete_data <- select_df[complete.cases(select_df$ss1.race, select_df$clust3_label), ]
table(complete_data$ss1.race, complete_data$clust3_label)

```


```{r}
colnames(df_wide_2)
```


```{r}
df_numeric %>%
  group_by(clust3_label) %>%
  summarize(mean = mean(ptg))
```
```{r}
colnames(df_numeric)
```



```{r}
summary(merge_model)
```
```{r}
# Coefficients for 'mid' and 'high' categories for the merge_model bad, possibly worse?
coefficients <- matrix(c(-2.655741, 0.5976946, 1.1773147, -0.6131220, -0.5464942, -0.3013413, 0.8617175, -1.9156269, -0.5266906,
                         -2.694907, 0.6851649, 0.8156827, -0.5029275, 0.2064163, -0.3846249, 0.9055494, -0.2315693, -0.2580464), 
                       ncol=9, byrow = TRUE)

# Standard errors for 'mid' and 'high' categories
std_errors <- matrix(c(0.9772299, 0.6027074, 0.3050416, 0.4096141, 0.7080132, 0.6792008, 0.6912819, 0.9299509, 0.8113750,
                       1.0457773, 0.6234894, 0.3241923, 0.4337231, 0.7711326, 0.7882847, 0.7785927, 0.8649438, 0.9160596), 
                     ncol=9, byrow = TRUE)

# Calculate odds ratios, z-values, and p-values
odds_ratios <- exp(coefficients)
z_values <- coefficients / std_errors
p_values <- 2 * pnorm(-abs(z_values))

# Variable names
variables <- c("Intercept", "wave1_etv", "wave1_pts", "Gender", "ss1.RaceBlack", "ss1.RaceHispanic", "ss1.RaceAsian", "ss1.RaceMixed", "ss1.RaceOther")

# Categories for each coefficient set
categories <- rep(c("mid", "high"), each = length(variables))

# Creating the results table
results_table <- data.frame(Category = categories,
                            Variable = rep(variables, 2),
                            Coefficients = c(coefficients),
                            Std.Errors = c(std_errors),
                            Odds.Ratios = c(odds_ratios),
                            Z.values = c(z_values),
                            P.values = round(c(p_values), 4))

# Displaying the results table
print(results_table)

```


```{r}
view(select_df)
```

```{r}
view(df_wide_2)
```
```{r}
data_merged <- merge(df_relabel, df_wide[c("SUBJID", "1")], by = "SUBJID", all.x = TRUE)
data_merged <- data_merged  %>% rename(clust3 = `1`)
```


```{r}
colnames(df_wide_2)
```

```{r}

df_relabel$SUBJID <- as.character(df_relabel$SUBJID)
df_wide$SUBJID <- as.character(df_wide$SUBJID)
data_merged <- data_merged[!is.na(data_merged$`1`), ]
view(data_merged)
```

##Note to self: df_numeric lost 6 entries, because they don't have an age for ss1.Age, so it's 194 instead of 200

```{r}
lcga3_linear$pprob

```

```{r}
view(df_wide)
```

```{r}
str(df_wide)
```

```{r}
data1 %>% filter(race_attrition1 == "1")
data1 %>% filter(SUBJID == "JC3017")
data1 %>% filter(SUBJID == "JC4027")
data1 %>% filter(SUBJID == "JC4149")
```

```{r}
df_numeric$clust3 <- as.factor(df_numeric$class)
df_numeric$class <- NULL
# Set class 3 (low) as the reference category
df_numeric$clust3 <- relevel(df_numeric$clust3, ref = "3")
```

```{r}
x <- multinom(clust3 ~ pts1 + etv1, data = df_numeric)
summary(x)
exp(coef(x))
```





```{r}
summary(df_numeric$clust3)
```

```{r}
data1 %>%
  group_by(ss1.Race) %>%
  summarise(
    mean_wave1_ptg = round(mean(wave1_ptg, na.rm = TRUE),3),
    sd_wave1_ptg = round(sd(wave1_ptg, na.rm = TRUE),3),
    
    mean_wave2_ptg = round(mean(wave2_ptg, na.rm = TRUE),3),
    sd_wave2_ptg = round(sd(wave2_ptg, na.rm = TRUE),3),
    
    mean_wave3_ptg = round(mean(wave3_ptg, na.rm = TRUE),3),
    sd_wave3_ptg = round(sd(wave3_ptg, na.rm = TRUE),3),
    
    mean_wave4_ptg = round(mean(wave4_ptg, na.rm = TRUE),3),
    sd_wave4_ptg = round(sd(wave4_ptg, na.rm = TRUE),3)
  )


```

```{r}
df_numeric <- df_numeric %>% 
  mutate(ptg_mean = mean(ptg))

df_numeric <- df_numeric %>% 
  mutate(ptg_sd = sd(ptg))

df_numeric <- df_numeric %>%
  mutate(Rating = case_when(
    ptg > (ptg_mean + ptg_sd) ~ "High PTG",
    ptg < (ptg_mean - ptg_sd) ~ "Low PTG",
    TRUE ~ "Medium PTG"))

cluster_himidlo <- table(df_numeric$Rating, df_numeric$clust3)
cluster_himidlo
chisq.test(cluster_himidlo)


df_numeric <- df_numeric %>% 
  mutate(etv_mean = mean(etv))

df_numeric <- df_numeric %>% 
  mutate(etv_sd = sd(etv))

df_numeric <- df_numeric %>%
  mutate(ETVRating = case_when(
    etv > (etv_mean + etv_sd) ~ "High etv",
    etv < (etv_mean - etv_sd) ~ "Low etv",
    TRUE ~ "Medium etv"))

etv_himidlo <- table(df_numeric$ETVRating, df_numeric$clust3)
etv_himidlo
chisq.test(etv_himidlo)

```





```{r}
lcga1_quad=hlme(ptg~wave+I(wave^2),
           subject="SUBJID",
           ng=1,
           data=df_numeric)
summary(lcga1_quad)

```
```{r}
lcga2_quad=gridsearch(rep=10,
                 maxiter=10,
                 minit=lcga1_quad,
                 hlme(ptg~wave+I(wave^2),
                      subject="SUBJID",
                      ng=2,
                      data=df_numeric,
                      mixture=~wave+I(wave^2)))
summary(lcga2_quad)
```
#Random Intercept - it's not as good as a fixed effects model
```{r}
lcga1_linear_RI = hlme(fixed = ptg ~ wave,
                    random = ~1,
                    subject = "SUBJID",
                    ng = 1,
                    data = df_numeric)
lcga2_linear_RI = gridsearch(rep=10,
                          maxiter=10,
                          minit=lcga1_linear_RI,  
                          hlme(ptg~wave,  
                               subject="SUBJID",
                               ng=2,
                               data=df_numeric,
                               mixture=~wave,
                               random=~1))  
lcga3_linear_RI = gridsearch(rep=10,
                          maxiter=10,
                          minit=lcga1_linear_RI,  
                          hlme(ptg~wave,  
                               subject="SUBJID",
                               ng=3,
                               data=df_numeric,
                               mixture=~wave,
                               random=~1))  
lcga4_linear_RI = gridsearch(rep=10,
                          maxiter=10,
                          minit=lcga1_linear_RI,  
                          hlme(ptg~wave,  
                               subject="SUBJID",
                               ng=4,
                               data=df_numeric,
                               mixture=~wave,
                               random=~1))  
lcga5_linear_RI = gridsearch(rep=10,
                          maxiter=10,
                          minit=lcga1_linear_RI,  
                          hlme(ptg~wave,  
                               subject="SUBJID",
                               ng=5,
                               data=df_numeric,
                               mixture=~wave,
                               random=~1))  
lcga6_linear_RI = gridsearch(rep=10,
                          maxiter=10,
                          minit=lcga1_linear_RI,  
                          hlme(ptg~wave,  
                               subject="SUBJID",
                               ng=6,
                               data=df_numeric,
                               mixture=~wave,
                               random=~1))  

```

##Linear model - fixed effects. Best fit
```{r}
lcga1_linear = hlme(fixed = ptg ~ wave,
                    subject = "SUBJID",
                    ng = 1,
                    data = df_numeric)
lcga2_linear = gridsearch(rep=10,
                          maxiter=10,
                          minit=lcga1_linear,  
                          hlme(ptg~wave,  
                               subject="SUBJID",
                               ng=2,
                               data=df_numeric,
                               mixture=~wave))  
lcga3_linear = gridsearch(rep=10,
                          maxiter=10,
                          minit=lcga1_linear,  
                          hlme(ptg~wave,  
                               subject="SUBJID",
                               ng=3,
                               data=df_numeric,
                               mixture=~wave))  
lcga4_linear = gridsearch(rep=10,
                          maxiter=10,
                          minit=lcga1_linear,  
                          hlme(ptg~wave,  
                               subject="SUBJID",
                               ng=4,
                               data=df_numeric,
                               mixture=~wave))  
lcga5_linear = gridsearch(rep=10,
                          maxiter=10,
                          minit=lcga1_linear,  
                          hlme(ptg~wave,  
                               subject="SUBJID",
                               ng=5,
                               data=df_numeric,
                               mixture=~wave))  
lcga6_linear = gridsearch(rep=10,
                          maxiter=10,
                          minit=lcga1_linear,  
                          hlme(ptg~wave,  
                               subject="SUBJID",
                               ng=6,
                               data=df_numeric,
                               mixture=~wave))  

```

##Linear model - fixed effects with demographic predictors
```{r}
lcga1_ctrls = hlme(fixed = ptg ~ wave + Gender + ss1.RaceWhite + ss1.RaceBlack + ss1.RaceHispanic + ss1.RaceAsian + ss1.RaceMixed + ss1.RaceOther,
                    subject = "SUBJID",
                    ng = 1,
                    data = df_numeric)
lcga2_ctrls = gridsearch(rep=10,
                          maxiter=10,
                          minit=lcga1_ctrls,  
                          hlme(ptg~wave + Gender + ss1.RaceWhite + ss1.RaceBlack + ss1.RaceHispanic + ss1.RaceAsian + ss1.RaceMixed + ss1.RaceOther,  
                               subject="SUBJID",
                               ng=2,
                               data=df_numeric,
                               mixture=~wave))  
lcga3_ctrls = gridsearch(rep=10,
                          maxiter=10,
                          minit=lcga1_ctrls,  
                          hlme(ptg~wave + Gender + ss1.RaceWhite + ss1.RaceBlack + ss1.RaceHispanic + ss1.RaceAsian + ss1.RaceMixed + ss1.RaceOther,  
                               subject="SUBJID",
                               ng=3,
                               data=df_numeric,
                               mixture=~wave))  
lcga4_ctrls = gridsearch(rep=10,
                          maxiter=10,
                          minit=lcga1_ctrls,  
                          hlme(ptg~wave + Gender + ss1.RaceWhite + ss1.RaceBlack + ss1.RaceHispanic + ss1.RaceAsian + ss1.RaceMixed + ss1.RaceOther,  
                               subject="SUBJID",
                               ng=4,
                               data=df_numeric,
                               mixture=~wave))  
lcga5_ctrls = gridsearch(rep=10,
                          maxiter=10,
                          minit=lcga1_ctrls,  
                          hlme(ptg~wave + Gender + ss1.RaceWhite + ss1.RaceBlack + ss1.RaceHispanic + ss1.RaceAsian + ss1.RaceMixed + ss1.RaceOther,  
                               subject="SUBJID",
                               ng=5,
                               data=df_numeric,
                               mixture=~wave))  
lcga6_ctrls = gridsearch(rep=10,
                          maxiter=10,
                          minit=lcga1_ctrls,  
                          hlme(ptg~wave + Gender + ss1.RaceWhite + ss1.RaceBlack + ss1.RaceHispanic + ss1.RaceAsian + ss1.RaceMixed + ss1.RaceOther,  
                               subject="SUBJID",
                               ng=6,
                               data=df_numeric,
                               mixture=~wave))  

```

```{r}
df_numeric %>% group_by(clust3) %>% 
  summarise(mean = mean(ptg))
```







```{r}
ggplot(df_numeric, aes(x=wave, y=ptg, colour = clust3, group = SUBJID)) +
  geom_point(size = .1) + geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = clust3)) +
  scale_color_manual(values=c( "blue","green","red"),
                        name = "Class",
                        breaks = c("mid","low","high"),
                        labels =c("Medium", "High", "Low")) +
    labs(
    title = "Three Class Model for Post-Traumatic Growth Trajectories",
    x = "Wave",
    y = "Post-Traumatic Growth"
  ) + 
  theme_minimal()
```
```{r}
ggplot(df_numeric, aes(x = wave, y = ptg, linetype = clust3, group = SUBJID)) +
  geom_point(size = .1) + 
  geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = clust3)) +
  scale_linetype_manual(values = c("solid", "dashed", "dotted"),
                        name = "Class",
                        breaks = c("mid","high","low"),
                        labels =c("Low", "High", "Medium")) +
    labs(
    title = "Three Class Model for Post-Traumatic Growth Trajectories",
    x = "Wave",
    y = "Post-Traumatic Growth"
  ) + 
  theme_minimal()

```


```{r}
ggplot(df_numeric, aes(x = wave, y = ptg, linetype = clust3, group = SUBJID)) +
  geom_point(size = .1) + 
  geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = clust3)) +
  scale_linetype_manual(values = c("solid", "dotted", "dashed"),
                        name = "Class",
                        breaks = c("mid", "low", "high"),
                        labels = c("Low", "Medium", "High")) +
  labs(
    title = "Three Class Model for Post-Traumatic Growth Trajectories",
    x = "Wave",
    y = "Post-Traumatic Growth"
  ) + 
  theme_minimal()

```
```{r}
ggplot(df_numeric, aes(x = wave, y = ptg, linetype = clust3, group = SUBJID)) +
  geom_point(size = .1) + 
  geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = clust3)) +
  scale_linetype_manual(values = c("solid", "dotted", "dashed"),
                        name = "Class",
                        breaks = c("high", "low", "mid"),
                        labels = c("High", "Medium", "Low")) +
  labs(
    title = "Three Class Model for Post-Traumatic Growth Trajectories",
    x = "Wave",
    y = "Post-Traumatic Growth"
  ) + 
  theme_minimal()

```


```{r}
pp_m3 <- postprob(lcga3_linear)
head(pp_m3)
```


```{r}
view(df_numeric)
```
```{r}
etv_group <- multinom(clust3 ~ etv1, data = df_numeric)
summary(etv_group)
exp(coef(etv_group))
```
```{r}
# For the `pts1` coefficient in class 2 - PTS at time 1 significantly increases the likelihood of being in the high class compared to the low class
coef_pts2 <- 0.7601352
se_pts2 <- 0.1464455
z_pts2 <- coef_pts2/se_pts2
p_value_pts2 <- 2 * (1 - pnorm(abs(z_pts2)))

p_value_pts2

coef_pts3 <- 0.9942172
se_pts3 <- 0.1364159
z_pts3 <- coef_pts3/se_pts3
p_value_pts3 <- 2 * (1 - pnorm(abs(z_pts3)))

p_value_pts3

coef_etv2 <- 1.821400
se_etv2 <- 0.3506944
z_etv2 <- coef_etv2/se_etv2
p_value_etv2 <- 2 * (1 - pnorm(abs(z_etv2)))

p_value_etv2

coef_etv3 <- 1.466475
se_etv3 <- 0.3306342
z_etv3 <- coef_etv3/se_etv3
p_value_etv3 <- 2 * (1 - pnorm(abs(z_etv3)))

p_value_etv3

```

#The log odds of being in the high group compared to the low group when ETV is 0 is -2.59; When ETV at time 1 is added, the log odds increase by 1.82 (OR = 6.18); Exposure to violence at Time 1 similarly affects the odds of being in the middle class versus the low class (-1.63; increased log odds to 1.47; OR = 4.33)

#Real change or benefit finding? 

```{r}
pts_group <- multinom(clust3 ~ pts1, data = df_numeric)
summary(pts_group)
exp(coef(pts_group))

pts_group_ctrls <- multinom(clust3 ~ pts1 + Gender + ss1.RaceBlack + ss1.RaceHispanic + ss1.RaceAsian + ss1.RaceMixed + ss1.RaceOther, data = df_numeric)
summary(pts_group_ctrls)
exp(coef(pts_group_ctrls))


etv_group_ctrls <- multinom(clust3 ~ etv1 + Gender + ss1.RaceBlack + ss1.RaceHispanic + ss1.RaceAsian + ss1.RaceMixed + ss1.RaceOther, data = df_numeric)
summary(etv_group_ctrls)
exp(coef(etv_group_ctrls))
df_numeric$clust3 <- factor(df_numeric$clust3, levels = c("3", "1", "2"))
fullmodel <- multinom(clust3 ~ etv1 + pts1 + Gender + ss1.RaceBlack + ss1.RaceHispanic + ss1.RaceAsian + ss1.RaceMixed + ss1.RaceOther, data = df_numeric)
fullmodel2 <- multinom(clust3_label ~ SUBJID + etv1 + pts1 + Gender + ss1.RaceBlack + ss1.RaceHispanic + ss1.RaceAsian + ss1.RaceMixed + ss1.RaceOther, data = df_numeric)
merged_df <- merged_df %>% 
  rename(clust3 = 1)
merge_model <- multinom(1 ~ wave1_etv + wave1_pts + Gender + ss1.RaceBlack + ss1.RaceHispanic + ss1.RaceAsian + ss1.RaceMixed + ss1.RaceOther, data = merged_df)
wide_model <- multinom(clust3_label ~ etv1 + pts1 + Gender + ss1.RaceBlack + ss1.RaceHispanic + ss1.RaceAsian + ss1.RaceMixed + ss1.RaceOther, data = df_wide_2)
summary(fullmodel)
summary(fullmodel2)
summary(wide_model)
exp(coef(fullmodel))
odds_ratios <- exp(coef(wide_model))
view(odds_ratios)
view(df_wide_2)
```

```{r}
#interpretable fullmodel with wave predictor
#pts and etv are non-sig but almost all race and gender predictors are? doesn't seem right

# Coefficients for categories 1 and 2 (assuming category 3 is the reference)
coefficients <- matrix(c(-2.850799, 0.01680523, 0.8302211, 1.1165364, -0.6633143, -0.68990742, -0.05595215, 0.8155134, -1.95982866, -0.4721150,
                         -3.463837, 0.03216293, 1.2181236, 0.7561324, -0.4879568, 0.07580734, -0.07301970, 1.0494525, 0.07627947, -0.3857532), 
                       ncol=10, byrow = TRUE)

# Standard errors for categories 1 and 2
std_errors <- matrix(c(0.6076433, 0.09435963, 0.3772647, 0.1665592, 0.2238906, 0.3818157, 0.3706941, 0.3742053, 0.5314823, 0.4322365,
                       0.6548465, 0.10044541, 0.3831468, 0.1776905, 0.2385173, 0.4297187, 0.4409313, 0.4320755, 0.4789691, 0.5132309), 
                     ncol=10, byrow = TRUE)

# Calculate odds ratios, z-values, and p-values
odds_ratios <- exp(coefficients)
z_values <- coefficients / std_errors
p_values <- 2 * pnorm(-abs(z_values))

# Variable names
variables <- c("Intercept", "wave", "etv1", "pts1", "Gender", "ss1.RaceBlack", "ss1.RaceHispanic", "ss1.RaceAsian", "ss1.RaceMixed", "ss1.RaceOther")

# Categories for each coefficient set
categories <- rep(c("1", "2"), each = length(variables))

# Creating the results table
results_table <- data.frame(Category = categories,
                            Variable = rep(variables, 2),
                            Coefficients = c(coefficients),
                            Std.Errors = c(std_errors),
                            Odds.Ratios = c(odds_ratios),
                            Z.values = c(z_values),
                            P.values = round(c(p_values), 4))

# Displaying the results table
print(results_table)

```

```{r}
length(merged_df$wave1_etv)
```

```{r}
view(merged_df)
```

#The log odds of being in the high group compared to the low group when PTS is 0 is -1.82; When PTS at time 1 is added, the log odds increase by 0.76 (OR = 2.14); PTS at Time 1 similarly affects the odds of being in the middle class versus the low class (-1.97; increased log odds to 0.99; OR = 2.70)

#Results are significant and align with pattern in exposure to violence

```{r}
# Coefficients for 'mid' and 'high' categories for df_wide_2 - the really bad results
coefficients <- matrix(c(-2.694907, -2.655741, 0.6851649, 0.5976946, 0.8156827, 1.1773147, 
                         -0.5029275, -0.6131220, 0.2064163, -0.5464942, -0.3846249, -0.3013413, 
                         0.9055494, 0.8617175, -0.2315693, -1.9156269, -0.2580464, -0.5266906), 
                       ncol=9, byrow = TRUE)

# Standard errors for 'mid' and 'high' categories
std_errors <- matrix(c(1.0457773, 0.9772299, 0.6234894, 0.6027074, 0.3241923, 0.3050416, 
                       0.4337231, 0.4096141, 0.7711326, 0.7080132, 0.7882847, 0.6792008, 
                       0.7785927, 0.6912819, 0.8649438, 0.9299509, 0.9160596, 0.8113750), 
                     ncol=9, byrow = TRUE)

# Calculate odds ratios, z-values, and p-values
odds_ratios <- exp(coefficients)
z_values <- coefficients / std_errors
p_values <- 2 * pnorm(-abs(z_values))

# Variable names
variables <- c("Intercept", "etv1", "pts1", "Gender", "ss1.RaceBlack", "ss1.RaceHispanic", "ss1.RaceAsian", "ss1.RaceMixed", "ss1.RaceOther")

# Categories for each coefficient set
categories <- rep(c("mid", "high"), each = length(variables))

# Creating the results table
results_table <- data.frame(Category = categories,
                            Variable = rep(variables, 2),
                            Coefficients = c(coefficients),
                            Std.Errors = c(std_errors),
                            Odds.Ratios = c(odds_ratios),
                            Z.values = c(z_values),
                            P.values = round(c(p_values), 4))

# Displaying the results table
print(results_table)
```

```{r}
# Coefficients for 'mid' and 'high' categories
coefficients <- matrix(c(-2.655741, 0.5976946, 1.1773147, -0.6131220, -0.5464942, -0.3013413, 0.8617175, -1.9156269, -0.5266906,
                         -2.694907, 0.6851649, 0.8156827, -0.5029275, 0.2064163, -0.3846249, 0.9055494, -0.2315693, -0.2580464), 
                       ncol=9, byrow = TRUE)

# Standard errors for 'mid' and 'high' categories
std_errors <- matrix(c(0.9772299, 0.6027074, 0.3050416, 0.4096141, 0.7080132, 0.6792008, 0.6912819, 0.9299509, 0.8113750,
                       1.0457773, 0.6234894, 0.3241923, 0.4337231, 0.7711326, 0.7882847, 0.7785927, 0.8649438, 0.9160596), 
                     ncol=9, byrow = TRUE)

# Calculate odds ratios, z-values, and p-values
odds_ratios <- exp(coefficients)
z_values <- coefficients / std_errors
p_values <- 2 * pnorm(-abs(z_values))

# Variable names
variables <- c("Intercept", "etv1", "pts1", "Gender", "ss1.RaceBlack", "ss1.RaceHispanic", "ss1.RaceAsian", "ss1.RaceMixed", "ss1.RaceOther")

# Categories for each coefficient set
categories <- rep(c("mid", "high"), each = length(variables))

# Creating the results table
results_table <- data.frame(Category = categories,
                            Variable = rep(variables, 2),
                            Coefficients = c(coefficients),
                            Std.Errors = c(std_errors),
                            Odds.Ratios = c(odds_ratios),
                            Z.values = c(z_values),
                            P.values = round(c(p_values), 4))

# Displaying the results table
print(results_table)

```

```{r}
data1 %>% group_by(ss1.RaceAsian) %>% 
  summarize(mean1 = mean(wave1_ptg), mean2 = mean(wave2_ptg, na.rm = TRUE), mean3 = mean(wave3_ptg, na.rm = TRUE), mean4 = mean(wave4_ptg, na.rm = TRUE))
```
```{r}
data1 %>% group_by(ss1.Sex) %>% 
  summarize(mean1 = mean(wave1_ptg), mean2 = mean(wave2_ptg, na.rm = TRUE), mean3 = mean(wave3_ptg, na.rm = TRUE), mean4 = mean(wave4_ptg, na.rm = TRUE))
```


```{r}
view(data_merged)
```
```{r}
df_wide_2 %>% group_by(clust3) %>% 
  summarize(mean = mean(ptg))
```


```{r}
# Assuming 'model' is your multinom model object --- results are from df_wide_2
# Extract coefficients and standard errors
coefficients <- matrix(c(2.65581991, -0.03895312, -0.59775991, 0.08741647, 
                         -1.1773087, -0.3616525, 0.6130970, 0.1101876, 
                         0.5465490, 0.7528523, 0.3013186, -0.0834360, 
                         -0.8617100, 0.0437572, 1.915692, 1.684007, 
                         0.5266890, 0.2686162), ncol=9, byrow = TRUE)

std_errors <- matrix(c(0.9772354, 0.9736890, 0.6027087, 0.5279210, 
                       0.3050418, 0.2735727, 0.4096147, 0.3841996, 
                       0.7080180, 0.7376744, 0.6792047, 0.7439356, 
                       0.6912869, 0.7010874, 0.9299593, 0.9475956, 
                       0.8113815, 0.8464008), ncol=9, byrow = TRUE)

# Calculate odds ratios, z-values, and p-values
odds_ratios <- exp(coefficients)
z_values <- coefficients / std_errors
p_values <- 2 * pnorm(-abs(z_values))

# Variable names (as per your model)
variables <- c("Intercept", "etv1", "pts1", "Gender", "ss1.RaceBlack", "ss1.RaceHispanic", "ss1.RaceAsian", "ss1.RaceMixed", "ss1.RaceOther")

# Creating the results table
results_table <- data.frame(Variable = rep(variables, 2),
                            Coefficients = c(coefficients),
                            Std.Errors = c(std_errors),
                            Odds.Ratios = c(odds_ratios),
                            Z.values = c(z_values),
                            P.values = round(c(p_values), 4))

# Displaying the results table
print(results_table)

```
```{r}
# Coefficients for 'mid' and 'high' categories
coefficients <- matrix(c(-2.850799, 0.01680523, 0.8302211, 1.1165364, -0.6633143, -0.68990742, -0.05595215, 0.8155134, -1.95982866, -0.4721150,
                         -3.463837, 0.03216293, 1.2181236, 0.7561324, -0.4879568, 0.07580734, -0.07301970, 1.0494525, 0.07627947, -0.3857532), 
                       ncol=10, byrow = TRUE)

# Standard errors for 'mid' and 'high' categories
std_errors <- matrix(c(0.6076433, 0.09435963, 0.3772647, 0.1665592, 0.2238906, 0.3818157, 0.3706941, 0.3742053, 0.5314823, 0.4322365,
                       0.6548465, 0.10044541, 0.3831468, 0.1776905, 0.2385173, 0.4297187, 0.4409313, 0.4320755, 0.4789691, 0.5132309), 
                     ncol=10, byrow = TRUE)

# Calculate odds ratios, z-values, and p-values
odds_ratios <- exp(coefficients)
z_values <- coefficients / std_errors
p_values <- 2 * pnorm(-abs(z_values))

# Variable names
variables <- c("Intercept", "wave", "etv1", "pts1", "Gender", "ss1.RaceBlack", "ss1.RaceHispanic", "ss1.RaceAsian", "ss1.RaceMixed", "ss1.RaceOther")

# Categories for each coefficient set
categories <- rep(c("mid", "high"), each = length(variables))

# Creating the results table
results_table <- data.frame(Category = categories,
                            Variable = rep(variables, 2),
                            Coefficients = c(coefficients),
                            Std.Errors = c(std_errors),
                            Odds.Ratios = c(odds_ratios),
                            Z.values = c(z_values),
                            P.values = round(c(p_values), 4))

# Displaying the results table
print(results_table)

```




```{r}


# Given coefficients and standard errors
coefficients_pts <- matrix(c(-0.2348698, 2.0192038,
                             -0.2820404, -1.2640003,
                             0.148799, 0.750763,
                             0.8316571, 0.5944501,
                             0.019784671, 0.004534868,
                             0.2466652, -0.8239916,
                             2.132344, 1.900284,
                             0.1065995, 0.4591572), ncol=2)

std_errors_pts <- matrix(c(0.4665645, 0.4229041,
                           0.1379694, 0.1565005,
                           0.2091711, 0.2209796,
                           0.4119177, 0.3799258,
                           0.4102430, 0.3710767,
                           0.3915013, 0.3754349,
                           0.5363947, 0.5272991,
                           0.4726280, 0.4334171), ncol=2)

variables_pts <- c("(Intercept)", "pts1", "Gender", "ss1.RaceBlack", "ss1.RaceHispanic", "ss1.RaceAsian", "ss1.RaceMixed", "ss1.RaceOther")
group <- rep(2:3, each=length(variables_pts))

z_values_pts <- coefficients_pts / std_errors_pts
p_values_pts <- 2 * (1 - pnorm(abs(z_values_pts)))
odds_ratios_pts <- exp(coefficients_pts)

# Create table
result_table_pts <- data.frame(Variable = rep(variables_pts, 2),
                               Group = group,
                               Coefficients = c(coefficients_pts),
                               Std.Errors = round(c(std_errors_pts), 3),
                               Z_values = c(z_values_pts),
                               P_values = round(c(p_values_pts), 3),
                               Odds_Ratios = c(odds_ratios_pts))

result_table_pts


```

```{r}
# Updated coefficients and standard errors
coefficients_wide <- matrix(c(2.65581991, -0.03895312, -0.59775991, 0.08741647, 
                              -1.1773087, -0.3616525, 0.6130970, 0.1101876, 
                              0.5465490, 0.7528523, 0.3013186, -0.0834360, 
                              -0.8617100, 0.0437572, 1.915692, 1.684007, 
                              0.5266890, 0.2686162), ncol=2)

std_errors_wide <- matrix(c(0.9772354, 0.9736890, 0.6027087, 0.5279210, 
                            0.3050418, 0.2735727, 0.4096147, 0.3841996, 
                            0.7080180, 0.7376744, 0.6792047, 0.7439356, 
                            0.6912869, 0.7010874, 0.9299593, 0.9475956, 
                            0.8113815, 0.8464008), ncol=2)

variables_wide <- c("(Intercept)", "etv1", "pts1", "Gender", "ss1.RaceBlack", "ss1.RaceHispanic", "ss1.RaceAsian", "ss1.RaceMixed", "ss1.RaceOther")
group <- rep(2:3, each=length(variables_wide))

# Calculating z-values, p-values, and odds ratios
z_values_wide <- coefficients_wide / std_errors_wide
p_values_wide <- 2 * (1 - pnorm(abs(z_values_wide)))
odds_ratios_wide <- exp(coefficients_wide)

# Creating the result table
result_table_wide <- data.frame(Variable = rep(variables_wide, 2),
                                Group = group,
                                Coefficients = c(coefficients_wide),
                                Std.Errors = round(c(std_errors_wide), 3),
                                Z_values = c(z_values_wide),
                                P_values = round(c(p_values_wide), 3),
                                Odds_Ratios = c(odds_ratios_wide))

# Displaying the result table
result_table_wide


```
```{r}
view(df_wide_2)
```



```{r}

# Calculate mean of 'ptg' for each 'clust3' group
mean_ptg_per_group <- df_numeric %>%
  group_by(clust3) %>%
  summarize(mean_ptg = mean(ptg, na.rm = TRUE))

# Calculate the relative frequency of each 'clust3' group
total_rows <- nrow(df_numeric)
relative_frequency <- df_numeric %>%
  count(clust3) %>%
  mutate(relative_freq = n / total_rows)

# Combine both results
combined_results <- left_join(mean_ptg_per_group, relative_frequency, by = "clust3")

combined_results


# Calculate mean of 'ptg' for each 'clust3' group
mean_ptg1_per_group <- df_numeric %>%
  group_by(clust3) %>%
  summarize(mean_ptg = mean(ptg1, na.rm = TRUE))

# Calculate the relative frequency of each 'clust3' group
total_rows1 <- nrow(df_numeric) %>% 
  filter(wave == 1)
relative_frequency <- df_numeric %>%
  count(clust3) %>%
  mutate(relative_freq = n / total_rows1)

# Combine both results
combined_results <- left_join(mean_ptg1_per_group, relative_frequency, by = "clust3")

combined_results

```

```{r}
# Filter df_numeric for wave1
df_numeric_wave1 <- df_numeric %>% 
  filter(wave == "1")  # Replace "wave1" with the correct value if needed

# Calculate total rows for wave1
total_rows1 <- nrow(df_numeric_wave1)

# Calculate mean of 'ptg' for each 'clust3' group in wave1
mean_ptg1_per_group <- df_numeric_wave1 %>%
  group_by(clust3) %>%
  summarize(mean_ptg = mean(ptg, na.rm = TRUE))

# Calculate the relative frequency of each 'clust3' group in wave1
relative_frequency <- df_numeric_wave1 %>%
  count(clust3) %>%
  mutate(relative_freq = n / total_rows1)

# Combine both results
combined_results <- left_join(mean_ptg1_per_group, relative_frequency, by = "clust3")

combined_results
```



```{r}
df_numeric %>% 
  group_by(clust3) %>% 
  summarize(mean = mean(ptg))
```
```{r}
# New coefficients and standard errors for the 'full' model
coefficients_full <- matrix(c(-2.809271, -3.385977,
                              0.8279136, 1.2139626,
                              1.1159326, 0.7560155,
                              -0.6614255, -0.4853164,
                              -0.68842963, 0.07733277,
                              -0.05566920, -0.07245934,
                              0.8153491, 1.0494506,
                              -1.95931152, 0.07750336,
                              -0.4707204, -0.3829401), ncol=2, byrow = TRUE)

std_errors_full <- matrix(c(0.5636177, 0.6072657,
                            0.3771898, 0.3829130,
                            0.1665366, 0.1776580,
                            0.2237136, 0.2383396,
                            0.3817878, 0.4297073,
                            0.3706973, 0.4409134,
                            0.3742127, 0.4320770,
                            0.5315079, 0.4789418,
                            0.4321425, 0.5130887), ncol=2, byrow = TRUE)

variables_full <- c("(Intercept)", "etv1", "pts1", "Gender", "ss1.RaceBlack", "ss1.RaceHispanic", "ss1.RaceAsian", "ss1.RaceMixed", "ss1.RaceOther")
group <- rep(2:3, each=length(variables_full))

z_values_full <- coefficients_full / std_errors_full
p_values_full <- 2 * (1 - pnorm(abs(z_values_full)))
odds_ratios_full <- exp(coefficients_full)

# Create table
result_table_full <- data.frame(Variable = rep(variables_full, 2),
                                Group = group,
                                Coefficients = c(coefficients_full),
                                Std.Errors = round(c(std_errors_full), 3),
                                Z_values = c(z_values_full),
                                P_values = round(c(p_values_full), 3),
                                Odds_Ratios = c(odds_ratios_full))

result_table_full

```
##Gender findings on ptg

```{r}
data1 %>% 
  group_by(ss1.Sex) %>% 
  summarize(
    mean1 = mean(wave1_ptg, na.rm = TRUE), 
    mean2 = mean(wave2_ptg, na.rm = TRUE), 
    mean3 = mean(wave3_ptg, na.rm = TRUE), 
    mean4 = mean(wave4_ptg, na.rm = TRUE)
  )
```
```{r}
df_numeric %>%
  group_by(clust3, Gender) %>%
  summarize(count = n())

```
```{r}
df_numeric %>%
  group_by(clust3) %>%
  summarize(mean_value = mean(ptg, na.rm = TRUE))

```


```{r}
summary(df_numeric$clust3)
```



```{r}
data1 %>% group_by(ss1.Sex) %>% 
  summarize(mean = mean(wave1_ptg))
```

```{r}
anova_clusters <- aov(pts~ clust3, data = df_numeric)
summary(anova_clusters)
```


```{r}
summary(lcga3_linear)
```

```{r}


# Given coefficients and standard errors for _etv
coefficients_etv <- matrix(c(-0.9314535, 2.0516078,
                             0.07587087, -1.81763289,
                             0.0399033, 0.1645796,
                             0.7685535, 0.5553363,
                             -0.08351517, -0.21353679,
                             0.2399803, -0.8565412,
                             2.000650, 1.674487,
                             -0.04319033, -0.04535837), ncol=2)

std_errors_etv <- matrix(c(0.5172626, 0.5411081,
                           0.2826796, 0.3505159,
                           0.2007665, 0.2010728,
                           0.4143215, 0.3608932,
                           0.4090419, 0.3469901,
                           0.3899148, 0.3541275,
                           0.5409396, 0.5139968,
                           0.4674401, 0.4029158), ncol=2)

variables_etv <- c("(Intercept)", "etv1", "Gender", "ss1.RaceBlack", "ss1.RaceHispanic", "ss1.RaceAsian", "ss1.RaceMixed", "ss1.RaceOther")
group_etv <- rep(2:3, each=length(variables_etv))

z_values_etv <- coefficients_etv / std_errors_etv
p_values_etv <- 2 * (1 - pnorm(abs(z_values_etv)))
odds_ratios_etv <- exp(coefficients_etv)

# Create table
result_table_etv <- data.frame(Variable = rep(variables_etv, 2),
                               Group = group_etv,
                               Coefficients = c(coefficients_etv),
                               Std.Errors = round(c(std_errors_etv),3),
                               Z_values = c(z_values_etv),
                               P_values = round(c(p_values_etv), 3),
                               Odds_Ratios = c(odds_ratios_etv))

result_table_etv

```




```{r}
result_table_etv
result_table_pts 
```

CLASS 1 - MID; CLASS 2 = HIGH, CLASS 3 = LOW




##LCMM with random intercept and random slope
```{r}
m1 <- lcmm::hlme(ptg ~ wave, random = ~wave, subject = "SUBJID", ng = 1, data = df_numeric)
m2 <- lcmm::hlme(ptg ~ wave, mixture = ~wave, random = ~wave, subject='SUBJID', ng = 2, data = df_numeric, B=m1)
m3 <- lcmm::hlme(ptg ~ wave, mixture = ~wave, random = ~wave, subject='SUBJID', ng = 3, data = df_numeric, B=m1)
m4 <- lcmm::hlme(ptg ~ wave, mixture = ~wave, random = ~wave, subject='SUBJID', ng = 4, data = df_numeric, B=m1)
m5 <- lcmm::hlme(ptg ~ wave, mixture = ~wave, random = ~wave, subject='SUBJID', ng = 5, data = df_numeric, B=m1)
m6 <- lcmm::hlme(ptg ~ wave, mixture = ~wave, random = ~wave, subject='SUBJID', ng = 6, data = df_numeric, B=m1)
```

#Table comparisons - random intercept and slope, random intercept, fixed
```{r}
tab1 <- as.data.frame(lcmm::summarytable(m1, m2, m3, m4, m5, m6))
tab2 <- as.data.frame(lcmm::summarytable(lcga1_linear_RI, lcga2_linear_RI, lcga3_linear_RI, lcga4_linear_RI, lcga5_linear_RI, lcga6_linear_RI))
tab3 <- as.data.frame(lcmm::summarytable(lcga1_linear, lcga2_linear, lcga3_linear, lcga4_linear, lcga5_linear, lcga6_linear))
tab4 <- as.data.frame(lcmm::summarytable(lcga1_ctrls, lcga2_ctrls, lcga3_ctrls, lcga4_ctrls, lcga5_ctrls, lcga6_ctrls))
```
```{r}
summary(lcga6_linear)
```
```{r}
summary(lcga6_ctrls)
```
```{r}
summary(lcga3_linear)
```
```{r}
conting_gender <- table(df_numeric$Gender, df_numeric$clust3)
conting_race <- table(df_numeric$ss1.Race, df_numeric$clust3)
conting_age <- table(df_numeric$ss1.Age, df_numeric$clust3)
chisq.test(conting_gender)
chisq.test(conting_race)
chisq.test(conting_age)
```

```{r}
conting_race
```

```{r}
chisq.posthoc.test(conting_race)
```






# 29.90% in Class 1; 27.32% in Class 2; 42.78% in Class 3 ; Class 1 = Mid; Class 2 = High; Class 3 = Low


```{r}
summary(df_numeric)
nrow(df_numeric)
view(data1)
table(df_numeric$wave)
```
```{r}
view(df_numeric)
```


```{r}
qplot(df_numeric$pts, df_numeric$ptg) + geom_smooth(method = "lm")
qplot(df_numeric$wave, df_numeric$ptg)
ggplot(df_numeric, aes(x = wave, y = ptg)) +
  geom_point() +
  geom_line()
linear_model <- lm(ptg ~ wave, data = df_numeric)
quadratic_model <- lm(ptg ~ wave + I(wave^2), data = df_numeric)
cubic_model <- lm(ptg ~ wave + I(wave*wave) + I(wave*wave*wave), data = df_numeric)
summary(linear_model)$r.squared
summary(quadratic_model)$r.squared
summary(cubic_model)$r.squared

AIC(linear_model)
AIC(quadratic_model)
AIC(cubic_model)

BIC(linear_model)
BIC(quadratic_model)
BIC(cubic_model)
par(mfrow=c(2,2))
plot(linear_model)
plot(quadratic_model)
plot(cubic_model)
```

```{r}
tab_cluster_gender <- table(df_numeric$clust3, df_numeric$Gender)
tab_cluster_race <- table(df_numeric$clust3, df_numeric$ss1.Race)
tab_cluster_age <- table(df_numeric$clust3, df_numeric$ss1.Age)
tab_cluster_missing <- table(df_numeric$clust3, df_numeric$category)
```


```{r}
chisq.test(tab_cluster_gender)
chisq.test(tab_cluster_missing)
chisq.test(tab_cluster_race) 
chisq.test(tab_cluster_age) 
t.test(pts1 ~ clust2, data = df_numeric)
t.test(etv1 ~ clust2, data = df_numeric)
aov1 <- aov(pts1 ~ clust3, data = df_numeric)
aov2 <- aov(etv1 ~ clust3, data = df_numeric)
summary(aov1)
summary(aov2)
```
```{r}
aov3 <- aov(wave1_etv ~ clust3_label_1, data = select_df)
summary(aov3)
aov4 <- aov(wave1_pts ~ clust3_label_1, data = select_df)
summary(aov4)
TukeyHSD(aov3)
TukeyHSD(aov4)
```
```{r}
colnames(select_df)
```

```{r}
view(df_numeric)
```


```{r}
select_df %>% group_by(clust3) %>% 
  summarize(meanpts = mean(wave1_pts), meanptg = mean(wave1_ptg), meanetv = mean(wave1_etv))
```
```{r}
colnames(select_df)
```


#FLAG Post hoc tests for pts1 and etv1  by cluster

```{r}
# Ensure the necessary package is installed and loaded
if (!require("multcomp")) install.packages("multcomp", dependencies = TRUE)
library(multcomp)

# Tukey HSD for aov1
tukey1 <- TukeyHSD(aov1)
print(tukey1)

# Tukey HSD for aov2
tukey2 <- TukeyHSD(aov2)
print(tukey2)

```



#FLAG = mean ptg pts1 and etv1 scores by cluster
```{r}
df_numeric %>% 
  group_by(clust3) %>% 
  summarize(mean = mean(pts1))

df_numeric %>% 
  group_by(clust3) %>% 
  summarize(mean = mean(etv1))


df_numeric %>% 
  group_by(clust3) %>% 
  summarize(mean = mean(ptg))
```




###CORRECT TO USE MISSING OTHERWISE TAKES ALL DATA FROM 1 - 3
```{r}
tab_category_gender <- table(missing$category, missing$Gender)
tab_category_race <- table(missing$category, missing$ss1.Race)
tab_category_age <- table(missing$category, missing$ss1.Age)
tab_category_cluster <- table(missing$category, missing$clust3)
chisq.test(tab_category_gender) 
chisq.test(tab_category_race)
chisq.test(tab_category_cluster)
t.test(pts1 ~ category, data = missing)
t.test(etv1 ~ category, data = missing)
t.test(ptg1 ~ category, data = missing)
```
```{r}
view(missing)
```

```{r}
tab_category_gender
```

```{r}
df_numeric$category
```


```{r}
table(df_numeric$category)
```


```{r}
tab_category_race
```
```{r}
chisq$expected
```

```{r}
summary(m1)
summary(m2)
```


```{r}
names(data1)
```


```{r}
# Create a contingency table
table_data <- with(data1, table(ss1.Race, category))

# Perform the chi-squared test
test_result <- chisq.test(table_data)

# Print the test result
print(test_result)

```
```{r}
table2 <- table(df_numeric$category, df_numeric$clust2)
chisq.test(table2)
```

```{r}
t.test(wave1_etv ~ category, data = data1)
```


```{r}
ptg_wave1 <- data1 %>% select(contains("ss1.ptg"))
pts_wave1 <- data1 %>% select(contains("ss1.pts"))
alpha(pts_wave1)
```
```{r}
view(pts_wave1)
```




```{r}
names(data)
```


```{r}
view(data)
```

Correlations
```{r}
corstars(df_var)
```
```{r}
correlations <- corstars(df_var)
```
```{r}
write.table(correlations, file = '/Users/kaylisealgrim/Library/CloudStorage/Box-Box/rutgers/projects/ptg/data/r_current/correlations_w4.txt', sep = ",") 
```

```{r}
pts_race <- aov(wave1_pts ~ ss1.Race, data = data1)
summary(pts_race)
```
```{r}
ptg_race <- aov(wave1_ptg ~ ss1.Race, data = data1)
summary(ptg_race)
```

```{r}
etv_race <- aov(wave1_etv ~ ss1.Race, data = data1)
summary(etv_race)
```
```{r}
pts_gender <- t.test(wave1_pts ~ Gender, data = data1)
summary(pts_gender)
```
```{r}
ptg_gender <- t.test(wave1_ptg ~ Gender, data = data1)
summary(ptg_gender)
```
```{r}
summary(lcga3_ctrls)
```



Data 1 Viewing
```{r}
class(data1)
dim(data1) 
names(data1) 
str(data1) 
glimpse(data1) 
summary(data1)
view(data1)
```




Descriptives and Visualizations

```{r}
ggplot(df_clean, aes(x = wave, y = ptg, group = SUBJID, color = factor(SUBJID))) +
  geom_line() +
  labs(x = "Wave", y = "PTG", color = "Participant ID") +
  theme_minimal() +
  theme(legend.position = "none")
```
```{r}
summary(df_var)
```


!!! Transform overweighted zeroes?

```{r}
ggplot(data,aes(x=wave1_ptg)) +   geom_histogram(fill = ("#a3c1ad"), binwidth = .4) + theme_minimal()
```
```{r}
ggplot(data,aes(x=wave2_ptg)) +   geom_histogram(fill = ("#a3c1ad"), binwidth = .4) + theme_minimal()
```
```{r}
ggplot(data,aes(x=wave3_ptg)) +   geom_histogram(fill = ("#a3c1ad"), binwidth = .4) + theme_minimal()
```
```{r}
ggplot(data,aes(x=wave4_ptg)) +   geom_histogram(fill = ("#a3c1ad"), binwidth = .4) + theme_minimal()
```
```{r}
ggplot(data,aes(x=wave1_pts)) +   geom_histogram(fill = ("#e0ac69"), binwidth = .2) + theme_minimal()
```
```{r}
ggplot(data,aes(x=wave1_etv)) +   geom_histogram(fill = ("#e76a6a"), binwidth = .2) + theme_minimal()
```
```{r}
ggplot(data,aes(x=wave1_gun)) +   geom_histogram(fill = ("#E6E6FA"), binwidth = .2) + theme_minimal()
```
```{r}
summary(df_var)
```

```{r}
ggplot(df_long, aes(x = wave, y = ptg, fill = wave)) +
  geom_boxplot() +
  scale_fill_manual(values = c("blue", "red", "yellow"))
```

```{r}
data1 %>% 
  summarise(mean = mean(ss1.Age, na.rm = TRUE), sd = sd(ss1.Age, na.rm = TRUE))
```
```{r}
data1 %>% 
  summarise(mean = mean(ss2.Age, na.rm = TRUE), sd = sd(ss2.Age, na.rm = TRUE))
```
```{r}
data1 %>% 
  summarise(mean = mean(ss3.Age, na.rm = TRUE), sd = sd(ss3.Age, na.rm = TRUE))
```
```{r}
data1 %>% 
  summarise(mean = mean(ss4.Age, na.rm = TRUE), sd = sd(ss4.Age, na.rm = TRUE))
```
```{r}
table(data1$ss1.Age)
table(data1$ss2.Age)
table(data1$ss3.Age)
```
```{r}
table(data1$ss1.Race)
table(data1$ss2.Race)
table(data1$ss3.Race)
```
```{r}
data1 %>%
  group_by(ss1.Race) %>%
  summarise(count = n()) %>%
  mutate(percentage = (count / sum(count)) * 100)

```
```{r}
data %>%
  filter(!is.na(ss2.Race)) %>%
  group_by(ss2.Race) %>%
  summarise(count = n()) %>%
  mutate(percentage = (count / sum(count)) * 100)
```
```{r}
data %>%
  filter(!is.na(ss3.Race)) %>%
  group_by(ss3.Race) %>%
  summarise(count = n()) %>%
  mutate(percentage = (count / sum(count)) * 100)
```
```{r}
data1 %>%
  filter(!is.na(ss4.Race)) %>%
  group_by(ss4.Race) %>%
  summarise(count = n()) %>%
  mutate(percentage = (count / sum(count)) * 100)
```
```{r}
cor.test(data1$wave4_ptg, data1$wave4_pts, na.rm = TRUE)
```


```{r}
data %>%
  filter(!is.na(ss3.Sex)) %>%
  group_by(ss3.Sex) %>%
  summarise(count = n()) %>%
  mutate(percentage = (count / sum(count)) * 100)
```
```{r}
data %>%
  filter(!is.na(ss4.Sex)) %>%
  group_by(ss4.Sex) %>%
  summarise(count = n()) %>%
  mutate(percentage = (count / sum(count)) * 100)
```
```{r}
data %>%
  filter(!is.na(ss2.Sex)) %>%
  group_by(ss2.Sex) %>%
  summarise(count = n()) %>%
  mutate(percentage = (count / sum(count)) * 100)
```


```{r}
ggplot(data1, aes(x = ss1.Race)) +
  geom_bar(fill = "pink", color = "black") +
  labs(x = "Race", y = "Count") +
  theme_minimal()
```


```{r}
df_clean %>%
  group_by(wave) %>%
  summarise(mean = mean(ptg), sd = sd(ptg))
```

```{r}
df_clean %>%
  group_by(wave) %>%
  summarise(mean = mean(etv), sd = sd(etv))
```

```{r}
df_clean %>%
  group_by(wave) %>%
  summarise(mean = mean(pts), sd = sd(pts))
```
```{r}
df_clean %>%
  group_by(wave) %>%
  summarise(mean = mean(gun), sd = sd(gun))
```

Latent Class Mixture Model
```{r}
m1 <- lcmm::hlme(ptg ~ wave, random = ~wave, subject = "SUBJID", ng = 1, data = df_numeric)
m2 <- lcmm::hlme(ptg ~ wave, mixture = ~wave, random = ~wave, subject='SUBJID', ng = 2, data = df_numeric, B=m1)
m3 <- lcmm::hlme(ptg ~ wave, mixture = ~wave, random = ~wave, subject='SUBJID', ng = 3, data = df_numeric, B=m1)
m4 <- lcmm::hlme(ptg ~ wave, mixture = ~wave, random = ~wave, subject='SUBJID', ng = 4, data = df_numeric, B=m1)
m5 <- lcmm::hlme(ptg ~ wave, mixture = ~wave, random = ~wave, subject='SUBJID', ng = 5, data = df_numeric, B=m1)
m6 <- lcmm::hlme(ptg ~ wave, mixture = ~wave, random = ~wave, subject='SUBJID', ng = 6, data = df_numeric, B=m1)
```

```{r}
# Model with no latent classes (for reference)
mq1 <- lcmm::hlme(ptg ~ wave + I(wave**2), random = ~wave, subject = "SUBJID", ng = 1, data = df_numeric)

# Models with 2 to 6 latent classes
mq2 <- lcmm::hlme(ptg ~ wave + I(wave**2), mixture = ~wave + I(wave**2), random = ~wave, subject = 'SUBJID', ng = 2, data = df_numeric, B=mq1)
mq3 <- lcmm::hlme(ptg ~ wave + I(wave**2), mixture = ~wave + I(wave**2), random = ~wave, subject = 'SUBJID', ng = 3, data = df_numeric, B=mq1)
mq4 <- lcmm::hlme(ptg ~ wave + I(wave**2), mixture = ~wave + I(wave**2), random = ~wave, subject = 'SUBJID', ng = 4, data = df_numeric, B=mq1)
mq5 <- lcmm::hlme(ptg ~ wave + I(wave**2), mixture = ~wave + I(wave**2), random = ~wave, subject = 'SUBJID', ng = 5, data = df_numeric, B=mq1)
mq6 <- lcmm::hlme(ptg ~ wave + I(wave**2), mixture = ~wave + I(wave**2), random = ~wave, subject = 'SUBJID', ng = 6, data = df_numeric, B=mq1)
```



```{r}
b <- ggplot(df_numeric, aes(x=wave, y=ptg, group=SUBJID, color=SUBJID)) + 
  geom_line() +
  geom_point() 

print(b)
```


```{r}
summary(m2)
```
```{r}
summary(m6)
```


```{r}
pp_m2 <- postprob(m2)
head(pp_m2)
```
```{r}
postprob(m3)
```


## The best fitting model appears to be a quadratic model with a single wave, but that's not necessarily what makes sense with our other intuitions and knowledge of this phenomenon. The second best fitting model is the linear 3 class, closely followed by the quadratic three class.
Model Selection
```{r}
#summary table for model comparison
tab1 <- as.data.frame(lcmm::summarytable(m1, m2, m3, m4, m5, m6))
tab2 <- as.data.frame(lcmm::summarytable(mq1, mq2, mq3, mq4, mq5, mq6))
tab3 <- as.data.frame(lcmm::summarytable(m1, m2, m3, m4, m5, m6, mq1, mq2, mq3, mq4, mq5, mq6))
```
```{r}
summary(m3)
```
```{r}
summary(mq3)
```

```{r}
tab1
```


```{r}
summary(m6)
```

```{r}
ggplot(tab2, aes(x=G, y=BIC)) +
  geom_point(size = 3, shape=21) + 
  geom_line(size = .3) +
  scale_y_continuous(limits = c(2100,2390)) +
  scale_x_continuous(breaks = c(1:6)) +
  xlab("Number of clusters") +
  theme_classic()
```

Predictors

```{r}
#Final models
m2etv_1 <- lcmm::hlme(ptg ~ wave + etv1, mixture = ~wave + etv1, random = ~wave, subject='SUBJID', classmb = ~ etv1, ng = 2, data = df_numeric)
m2pts_1 <- lcmm::hlme(ptg ~ wave + pts1, mixture = ~wave + pts1, random = ~wave, subject='SUBJID', classmb = ~ pts1, ng = 2, data = df_numeric)
m2etv_full <- lcmm::hlme(ptg ~ wave + etv1 + ss1.RaceBlack + ss1.RaceHispanic + ss1.RaceAsian + ss1.RaceMixed + ss1.RaceOther, mixture = ~wave + etv1 + ss1.RaceBlack + ss1.RaceHispanic + 
                           ss1.RaceAsian + ss1.RaceMixed + ss1.RaceOther, random = ~wave, subject='SUBJID', classmb = ~ etv1, ng = 2, data = df_numeric)
m2pts_full <- lcmm::hlme(ptg ~ wave + pts1 + ss1.RaceBlack + ss1.RaceHispanic + ss1.RaceAsian + ss1.RaceMixed + ss1.RaceOther, mixture = ~wave + pts1 + ss1.RaceBlack + ss1.RaceHispanic + ss1.RaceAsian + ss1.RaceMixed + ss1.RaceOther, random = ~wave, subject='SUBJID', classmb = ~ pts1, ng = 2, data = df_numeric)
m2race_full <- lcmm::hlme(ptg ~ wave + ss1.RaceBlack + ss1.RaceHispanic + ss1.RaceAsian + ss1.RaceMixed + ss1.RaceOther, mixture = ~wave + ss1.RaceBlack + ss1.RaceHispanic + ss1.RaceAsian + ss1.RaceMixed + ss1.RaceOther, random = ~wave, subject='SUBJID', ng = 2, data = df_numeric)
```



```{r}
summary(m2)
summary(m2etv_1)
summary(m2pts_1)
summary(m2pts_full)
summary(m2etv_full)
summary(m2race_full)
```


```{r}
summary(m2gender)
```
```{r}
summary(m2race)
```

```{r}
summary(m2age)
```









```{r}
data1 %>% 
  summarise(mean1 = mean(wave1_etv), mean2 = mean(wave2_etv, na.rm = TRUE), mean3 = mean(wave3_etv, na.rm = TRUE))
```
```{r}
data1 %>% 
  summarise(mean1 = mean(wave1_gun), mean2 = mean(wave2_gun, na.rm = TRUE), mean3 = mean(wave3_gun, na.rm = TRUE))
```

```{r}
d <- aov(etv ~ wave, data = df_numeric)
```
```{r}
summary(d)
```

```{r}
e <- aov(gun ~ wave, data = df_numeric)
```
```{r}
summary(e)
```
```{r}
data1 %>% 
  summarise(mean1 = mean(wave1_pts), mean2 = mean(wave2_pts, na.rm = TRUE), mean3 = mean(wave3_pts, na.rm = TRUE))
```

```{r}
f <- aov(pts ~ wave, data = df_numeric)
```
```{r}
summary(f)
```


```{r}
summary(m3)
```
```{r}
summary(m2)
```


```{r}
table(m3$pprob[,2])
```
```{r}
table(m2$pprob[,2])
```
```{r}
table(mq3$pprob[,2])
```
```{r}
table(mq2$pprob[,2])
```
INTRODUCE CLUSTERS BELOW

```{r}
df_numeric <- merge(df_numeric, round(m3$pprob, 3)[, 1:2], by = "SUBJID")
df_numeric$clust3 <- as.factor(df_numeric$class)
df_numeric$class <- NULL
df_numeric <- merge(df_numeric, round(m2$pprob, 3)[, 1:2], by = "SUBJID")
df_numeric$clust2 <- as.factor(df_numeric$class)
df_numeric$class <- NULL
df_numeric <- merge(df_numeric, round(mq3$pprob, 3)[, 1:2], by = "SUBJID")
df_numeric$clustq3 <- as.factor(df_numeric$class)
df_numeric$class <- NULL
df_numeric <- merge(df_numeric, round(mq2$pprob, 3)[, 1:2], by = "SUBJID")
df_numeric$clustq2 <- as.factor(df_numeric$class)
df_numeric$class <- NULL
```

```{r}
summary(m1)
```



Post-traumatic stress is significantly higher for those in Group 2

```{r}
conting_race
```


```{r}
t.test(pts3 ~ clust2, data = df_numeric)
```

```{r}
t.test(pts1 ~ clust2, data = df_numeric)
```
```{r}
t.test(etv1 ~ clust2, data = df_numeric)
```
Post-traumatic stress at time 3 was also higher for the high increasing post-traumatic growth group

Cluster membership does not significantly differ by gender, age,  or race.

```{r}
ggplot(df_numeric, aes(x=wave, y=ptg, colour = etvclust2, group = SUBJID)) +
  geom_point(size = .1) + geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = etvclust2)) +
  scale_color_manual(values=c( "red","green"))
```
```{r}
ggplot(df_numeric, aes(x=wave, y=ptg, colour = ptsclust2, group = SUBJID)) +
  geom_point(size = .1) + geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = ptsclust2)) +
  scale_color_manual(values=c( "red","green"))
```
```{r}
ggplot(df_numeric, aes(x=wave, y=ptg, colour = clust2, group = SUBJID)) +
  geom_point(size = .1) + geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = clust2)) +
  scale_color_manual(values=c( "red","green"))
```
```{r}
ggplot(df_numeric, aes(x=wave, y=ptg, colour = clust3, group = SUBJID)) +
  geom_point(size = .1) + geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = clust3)) +
  scale_color_manual(values=c( "red","green","blue"))
```

```{r}
df_numeric$clust2
```

Cleaned table
```{r}
library(ggplot2)

ggplot(df_numeric, aes(x=wave, y=ptg, colour = clust2, group = SUBJID)) +
  geom_point(size = .1) +
  geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = clust2)) +
  scale_color_manual(values=c("red","green"), 
                     name = "Class", 
                     breaks = c("1", "2"), 
                     labels = c("Low-Decreasing", "High-Increasing")) +
  labs(
    title = "Two Class Model for Post-Traumatic Growth Trajectories",
    x = "Wave",
    y = "Post-Traumatic Growth"
  ) + 
  theme_minimal()

```
```{r}
ggplot(df_numeric, aes(x=wave, y=ptg, linetype = clustq3, group = SUBJID)) +
  geom_point(size = .1) +
  geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = clust2)) +
  scale_linetype_manual(values=c("solid","dashed"), 
                        name = "Class", 
                        breaks = c("1", "2"), 
                        labels = c("Low-Decreasing", "High-Increasing")) +
  labs(
    title = "Two Class Model for Post-Traumatic Growth Trajectories",
    x = "Wave",
    y = "Post-Traumatic Growth"
  ) + 
  theme_minimal()

ggplot(df_numeric, aes(x=wave, y=ptg, linetype = clustq3, group = SUBJID)) +
  geom_point(size = .1) + 
  geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = clustq3)) +
  scale_color_manual(values=c( "red","green","blue"))

ggplot(df_numeric, aes(x=wave, y=ptg, linetype = clust3, group = SUBJID)) +
  geom_point(size = .1) + 
  geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = clustq3)) +
  scale_color_manual(values=c( "red","green","blue"))
```
```{r}
ggplot(df_numeric, aes(x=wave, y=ptg, linetype = clust2, group = SUBJID)) +
  geom_point(size = .1) +
  geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = clust2)) +
  scale_linetype_manual(values=c("solid","dashed"), 
                        name = "Class", 
                        breaks = c("1", "2"), 
                        labels = c("Low-Decreasing", "High-Increasing")) +
  labs(
    title = "Two Class Model for Post-Traumatic Growth Trajectories",
    x = "Wave",
    y = "Post-Traumatic Growth"
  ) + 
  theme_minimal()
```


```{r}
ggplot(df_numeric, aes(x=wave, y=ptg, colour = etvclust2, group = SUBJID)) +
  geom_point(size = .1) +
  geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = etvclust2)) +
  scale_color_manual(values=c("red","green"), 
                     name = "Class", 
                     breaks = c("1", "2"), 
                     labels = c("Low-Decreasing", "High-Increasing")) +
  labs(
    title = "Two Class Model for Post-Traumatic Growth Trajectories with ETV Covariate",
    x = "Wave",
    y = "Post-Traumatic Growth"
  ) + 
  theme_minimal()

```

```{r}
library(ggplot2)

ggplot(df_numeric, aes(x=wave, y=ptg, colour = ptsclust2, group = SUBJID)) +
  geom_point(size = .1) +
  geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = ptsclust2)) +
  scale_color_manual(values=c("red","green"), 
                     name = "Class", 
                     breaks = c("1", "2"), 
                     labels = c("Low-Decreasing", "High-Increasing")) +
  labs(
    title = "Two Class Model for Post-Traumatic Growth Trajectories with PTS Covariate",
    x = "Wave",
    y = "Post-Traumatic Growth"
  ) + 
  theme_minimal()

```

```{r}
summary(m2)
```


Age and Gender Differences

```{r}
t.test(ptg ~ Gender, data = df_numeric)
```
```{r}
unique(df_numeric$Gender)
```


Models
```{r}
#intercept only model
baseline_w1 <- lmer(ptg ~ 1 + (1 | SUBJID), data = df_merged, REML = FALSE)
```
```{r}
fixed_pts_wave1 <- lmer(ptg ~ 1 + pts + (1 | SUBJID), data = df_merged, REML = FALSE)
```
```{r}
#random fixed pts effect
ranfixed_pts_wave1 <- lmer(ptg ~ 1 + pts + (1 + pts | SUBJID), data = df_merged, REML = FALSE)
#getting result boundary (singular) fit: see ?isSingular
```
```{r}
#fixed etv effect
fixed_etv_wave1 <- lmer(ptg ~ 1 + etv + (1 | SUBJID), data = df_merged, REML = FALSE)
```
```{r}
#random fixed etv effect
ranfixed_etv_wave1 <- lmer(ptg ~ 1 + etv + (1 + etv | SUBJID), data = df_merged, REML = FALSE)
```

```{r}
etv_mod_c <- anova(baseline_w1, ranfixed_etv_wave1, fixed_etv_wave1)
```
```{r}
print(etv_mod_c)
```
```{r}
pts_mod_c <- anova(baseline_w1, ranfixed_pts_wave1, fixed_pts_wave1)
```
```{r}
print(pts_mod_c)
```


```{r}
summary(baseline_w1)
```

```{r}
summary(fixed_pts_wave1)
```
With PTS at 0, the average PTG is 2.55. Every time PTS goes up by a point, the average PTG score rises by .38. The standard deviation for the residual dropped a very small amount.

```{r}
summary(ranfixed_pts_wave1)
```
```{r}
summary(fixed_etv_wave1)
```

```{r}
summary(ranfixed_etv_wave1)
```
```{r}
#intercept only model
baseline <- lmer(ptg ~ 1 + (1 | SUBJID), data = df_clean)
```
```{r}
#fixed pts effect
fixed_pts <- lmer(ptg ~ 1 + pts + (1 | SUBJID), data = df_clean)
```
```{r}
#random fixed pts effect
ranfixed_pts <- lmer(ptg ~ 1 + pts + (1 + pts | SUBJID), data = df_clean)
```

```{r}
#fixed etv effect
fixed_etv <- lmer(ptg ~ 1 + etv + (1 | SUBJID), data = df_clean)
```
```{r}
#random fixed etv effect
ranfixed_etv <- lmer(ptg ~ 1 + etv + (1 + etv | SUBJID), data = df_clean)
```

OKAY! The random and fixed effects model for PTS and ETV offers the best fit. 

```{r}
summary(baseline)
```
The mean PTG is 3.44.
```{r}
summary(fixed_pts)
```
With PTS at 0, the average PTG is 2.55. Every time PTS goes up by a point, the average PTG score rises by .38. The standard deviation for the residual dropped a very small amount.

```{r}
summary(ranfixed_pts)
```
```{r}
summary(fixed_etv)
```

```{r}
summary(ranfixed_etv)
```
For ETV, a fixed effect model appears worse than baseline, but a linear mixed effect model with random effects improves on the model.

```{r}
summary(baseline)
```
The mean PTG is 3.44.




```{r}
data3 <- na.omit(data.frame(data1$wave1_ptg, data1$wave2_ptg, data1$wave3_ptg))
```
```{r}
ptg.cov <- cov(data3)
```
```{r}
data3
```

```{r}
pc1_ptg <- lm(wave2_ptg ~ wave1_ptg + wave1_pts, data = data1)
pc2_ptg <- lm(wave3_ptg ~ wave2_ptg + wave2_pts, data = data1)
pc1_pts <- lm(wave2_pts ~ wave1_ptg + wave1_pts, data = data1)
pc2_pts <- lm(wave3_pts ~ wave2_ptg + wave2_pts, data = data1)
```
```{r}
summary(pc1_pts)
```




3. Latent class mixture model



```{r}
ggplot(data = df_numeric, aes(x = pts, y = ptg, color = factor(wave))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "PTS", y = "PTG", color = "Wave") +
  theme_minimal()
```









