---
title: "4 Wave PTG Latent Growth Curves and Latent Class Mixture Models"
output:
  word_document: default
  html_notebook: default
  pdf_document: default
---
Summary: We first (1) tested the latent growth curve model for the entire dataset. We then (2) assessed 1 - 6 latent classes for the model, testing both quadratic and linear models. The best fit was the single trajectory model, but the three trajectory model was very close. As the models were comparable, the linear model with fewer parameters was selected over the quadratic model. (4) Using these models as outcomes, we predicted group membership from time 1 exposure to violence and time 1 post-traumatic stress using multinomial logistic regression.

#dataframe names
data - complete dataset
data1 - variables of interest for all waves
data2 - variables of interest with wave 1 predictors
df_long - variables of interest in long format
df_clean - long format dataset with NAs removed and high - low categories created for post-traumatic growth
df_pred - pts and etv at time 1
df_numeric - df_clean with SUBJID in numeric form to analyze

#loaded packages
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(knitr)
library(lavaan)
library(lcmm)
library(lme4)
corstars <-function(x, method=c("pearson", "spearman"), removeTriangle=c("upper", "lower"),
                     result=c("none", "html", "latex")){
    #Compute correlation matrix
    require(Hmisc)
    x <- as.matrix(x)
    correlation_matrix<-rcorr(x, type=method[1])
    R <- correlation_matrix$r # Matrix of correlation coeficients
    p <- correlation_matrix$P # Matrix of p-value 
    
    ## Define notions for significance levels; spacing is important.
    mystars <- ifelse(p < .0001, "****", ifelse(p < .001, "*** ", ifelse(p < .01, "**  ", ifelse(p < .05, "*   ", "    "))))
    
    ## trunctuate the correlation matrix to two decimal
    R <- format(round(cbind(rep(-1.11, ncol(x)), R), 2))[,-1]
    
    ## build a new matrix that includes the correlations with their apropriate stars
    Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x))
    diag(Rnew) <- paste(diag(R), " ", sep="")
    rownames(Rnew) <- colnames(x)
    colnames(Rnew) <- paste(colnames(x), "", sep="")
    
    ## remove upper triangle of correlation matrix
    if(removeTriangle[1]=="upper"){
      Rnew <- as.matrix(Rnew)
      Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
      Rnew <- as.data.frame(Rnew)
    }
    
    ## remove lower triangle of correlation matrix
    else if(removeTriangle[1]=="lower"){
      Rnew <- as.matrix(Rnew)
      Rnew[lower.tri(Rnew, diag = TRUE)] <- ""
      Rnew <- as.data.frame(Rnew)
    }
    
    ## remove last column and return the correlation matrix
    Rnew <- cbind(Rnew[1:length(Rnew)-1])
    if (result[1]=="none") return(Rnew)
    else{
      if(result[1]=="html") print(xtable(Rnew), type="html")
      else print(xtable(Rnew), type="latex") 
    }
} 
```

#all initial dataframes and recoding
```{r}
data <- read.csv("ptg_data.csv")
data <- data %>% 
  mutate(wave4_pts = rowMeans(select(.,ss4.pts01,ss4.pts02,ss4.pts03,ss4.pts04,ss4.pts05,ss4.pts06,ss4.pts07,ss4.pts08,ss4.pts09)))
data <- data %>% 
  mutate(wave4_ptg = rowMeans(select(.,ss4.ptg01,ss4.ptg02,ss4.ptg03,ss4.ptg04,ss4.ptg05,ss4.ptg06,ss4.ptg07,ss4.ptg08,ss4.ptg09,ss4.ptg10)))
data <- data %>% 
  mutate(wave4_etv = rowMeans(select(.,ss4.NVL01,ss4.NVL02,ss4.NVL03,ss4.NVL04,ss4.NVL05,ss4.NVL06,ss4.NVL07,ss4.NVL08,ss4.NVL09,ss4.NVL10,ss4.NVL11, ss4.NVL12,ss4.NVL13,ss4.NVL14,ss4.NVL15,ss4.NVL16,ss4.NVL17,ss4.NVL18,ss4.NVL19,ss4.NVL20)))
data1 <- data %>% select(c('SUBJID','ss1.Age','ss1.Sex','ss1.Race','ss2.Age','ss2.Sex','ss2.Race','ss3.Age','ss3.Sex','ss3.Race','ss4.Age','ss4.Sex','ss4.Race','wave1_ptg','wave2_ptg','wave3_ptg','wave4_ptg','wave1_pts','wave2_pts','wave3_pts','wave4_pts','wave1_etv','wave2_etv','wave3_etv','wave4_etv'))
##getting natural log for relevant variables
data1 <- data1 %>%
  mutate(across(matches("^wave\\d+_ptg$"), list(logptg = ~log(.))))
data1 <- data1 %>%
rename(wave1_logptg = wave1_ptg_logptg)
data1 <- data1 %>%
rename(wave2_logptg = wave2_ptg_logptg)
data1 <- data1 %>%
rename(wave3_logptg = wave3_ptg_logptg)
data1 <- data1 %>%
rename(wave4_logptg = wave4_ptg_logptg)


data1$ss1.Race <- factor(data1$ss1.Race,
                         levels = c(1,2,3,4,5,6),
                         labels = c("Black", "Hispanic", "Asian","White","Mixed","Other"))
data1$ss2.Race <- factor(data1$ss2.Race,
                         levels = c(1,2,3,4,5,6),
                         labels = c("Black", "Hispanic", "Asian","White","Mixed","Other"))
data1$ss3.Race <- factor(data1$ss3.Race,
                         levels = c(1,2,3,4,5,6),
                         labels = c("Black", "Hispanic", "Asian","White","Mixed","Other"))
data1$ss4.Race <- factor(data1$ss4.Race,
                         levels = c(1,2,3,4,5,6),
                         labels = c("Black", "Hispanic", "Asian","White","Mixed","Other"))
data1 <- data1 %>% mutate(ss1.Sex=recode(ss1.Sex,'1'='1','2'='0'))
data1 <- data1 %>% mutate(ss1.Age=as.numeric(ss1.Age), na.rm = TRUE)
data1 <- data1 %>% mutate(ss2.Age=as.numeric(ss2.Age), na.rm = TRUE)
data1 <- data1 %>% mutate(ss3.Age=as.numeric(ss3.Age), na.rm = TRUE)
data1 <- data1 %>% mutate(Gender=as.numeric(ss1.Sex), na.rm = TRUE)
#create category variable for whether or not they completed they entire study
data1 <- data1 %>%
  mutate(category = ifelse(is.na(ss4.Sex), "No Data", "Has Data"))
# Relevel the factor to make '4' the reference category
data1$ss1.Race <- relevel(factor(data1$ss1.Race), ref = 'White')
mat <- model.matrix(~ ss1.Race - 1, data = data1)
df_dummies <- as.data.frame(mat)
data1 <- cbind(data1, df_dummies)
df_long <- data1 %>%
  pivot_longer(
    cols = -c(SUBJID, `ss1.Age`, `Gender`, `ss1.Race`,`category`,`ss1.RaceWhite`,`ss1.RaceBlack`,`ss1.RaceHispanic`,
              `ss1.RaceAsian`,`ss1.RaceMixed`,`ss1.RaceOther`), # Exclude SUBJID from reshaping
    names_to = c("wave", ".value"),  # Specify where parts of the old column names will go
    names_pattern = "(wave\\d+)_(ptg|pts|etv|logptg)"  # Use a regular expression to match old column names
  )
df_clean <- na.omit(df_long)
# Remove the letters from the 'SUBJID' column + wave column
df_clean$SUBJID <- as.numeric(gsub("[A-Za-z]", "", df_clean$SUBJID))
df_clean$wave <- as.numeric(sub("wave", "", df_clean$wave))
df_numeric <- df_clean
#create df for correlation table
df_var <- data1 %>% 
  select("wave1_ptg","wave2_ptg","wave3_ptg","wave4_ptg","wave1_pts","wave2_pts","wave3_pts","wave4_pts","wave1_etv","wave2_etv","wave3_etv","wave4_etv")
df_numeric <- df_numeric %>%
  group_by(SUBJID) %>%
  mutate(etv1 = first(etv), pts1 = first(pts), ptg1 = first(ptg)) %>%
  ungroup()
```
```{r}
unique(df_numeric$wave)
```


#inspecting the data
```{r}
class(data1)
dim(data1) 
names(data1) 
str(data1) 
glimpse(data1) 
summary(data1)
```

#means and standard deviations
```{r}
df_clean %>%
  group_by(wave) %>%
  summarise(mean = mean(ptg), sd = sd(ptg))
```

```{r}
df_clean %>%
  group_by(wave) %>%
  summarise(mean = mean(etv), sd = sd(etv))
```

```{r}
df_clean %>%
  group_by(wave) %>%
  summarise(mean = mean(pts), sd = sd(pts))
```

#bivariate correlations
```{r}
correlations <- corstars(df_var)
```

#variable distributions
```{r}
ggplot(data,aes(x=wave1_ptg)) +   geom_histogram(fill = ("#a3c1ad"), binwidth = .4) + theme_minimal()
ggplot(data,aes(x=wave2_ptg)) +   geom_histogram(fill = ("#a3c1ad"), binwidth = .4) + theme_minimal()
ggplot(data,aes(x=wave3_ptg)) +   geom_histogram(fill = ("#a3c1ad"), binwidth = .4) + theme_minimal()
ggplot(data,aes(x=wave4_ptg)) +   geom_histogram(fill = ("#a3c1ad"), binwidth = .4) + theme_minimal()
ggplot(data,aes(x=wave1_pts)) +   geom_histogram(fill = ("#e0ac69"), binwidth = .2) + theme_minimal()
ggplot(data,aes(x=wave1_etv)) +   geom_histogram(fill = ("#e76a6a"), binwidth = .2) + theme_minimal()
```
#ptg trajectories for all participants by wave
```{r}
ggplot(df_clean, aes(x = wave, y = ptg, group = SUBJID, color = factor(SUBJID))) +
  geom_line() +
  labs(x = "Wave", y = "PTG", color = "Participant ID") +
  theme_minimal() +
  theme(legend.position = "none")
```

#descriptives and reliability
##race
```{r}
ggplot(data1, aes(x = ss1.Race)) +
  geom_bar(fill = "pink", color = "black") +
  labs(x = "Race", y = "Count") +
  theme_minimal()
data1 %>%
  group_by(ss1.Race) %>%
  summarise(count = n()) %>%
  mutate(percentage = (count / sum(count)) * 100)
data1 %>%
  filter(!is.na(ss2.Race)) %>%
  group_by(ss2.Race) %>%
  summarise(count = n()) %>%
  mutate(percentage = (count / sum(count)) * 100)
data1 %>%
  filter(!is.na(ss3.Race)) %>%
  group_by(ss3.Race) %>%
  summarise(count = n()) %>%
  mutate(percentage = (count / sum(count)) * 100)
data1 %>%
  filter(!is.na(ss4.Race)) %>%
  group_by(ss4.Race) %>%
  summarise(count = n()) %>%
  mutate(percentage = (count / sum(count)) * 100)
```
##age
```{r}
table(data1$ss1.Age)
table(data1$ss2.Age)
table(data1$ss3.Age)
table(data1$ss4.Age)

data1 %>% 
  summarise(mean = mean(ss1.Age, na.rm = TRUE), sd = sd(ss1.Age, na.rm = TRUE))
data1 %>% 
  summarise(mean = mean(ss2.Age, na.rm = TRUE), sd = sd(ss2.Age, na.rm = TRUE))
data1 %>% 
  summarise(mean = mean(ss3.Age, na.rm = TRUE), sd = sd(ss3.Age, na.rm = TRUE))
data1 %>% 
  summarise(mean = mean(ss4.Age, na.rm = TRUE), sd = sd(ss4.Age, na.rm = TRUE))
```

```{r}
tab_cluster_gender <- table(df_numeric$clust2, df_numeric$Gender)
tab_cluster_race <- table(df_numeric$clust2, df_numeric$ss1.Race)
tab_cluster_age <- table(df_numeric$clust2, df_numeric$ss1.Age)
```
```{r}
chisq.test(tab_cluster_gender)
chisq.test(tab_cluster_race) 
chisq.test(tab_cluster_age) 
t.test(pts1 ~ clust2, data = df_numeric)
t.test(etv1 ~ clust2, data = df_numeric)
```


#selection based on number of classes
```{r}
m1 <- lcmm::hlme(ptg ~ wave, random = ~wave, subject = "SUBJID", ng = 1, data = df_numeric)
m2 <- lcmm::hlme(ptg ~ wave, mixture = ~wave, random = ~wave, subject='SUBJID', ng = 2, data = df_numeric, B=m1)
m3 <- lcmm::hlme(ptg ~ wave, mixture = ~wave, random = ~wave, subject='SUBJID', ng = 3, data = df_numeric, B=m1)
m4 <- lcmm::hlme(ptg ~ wave, mixture = ~wave, random = ~wave, subject='SUBJID', ng = 4, data = df_numeric, B=m1)
m5 <- lcmm::hlme(ptg ~ wave, mixture = ~wave, random = ~wave, subject='SUBJID', ng = 5, data = df_numeric, B=m1)
m6 <- lcmm::hlme(ptg ~ wave, mixture = ~wave, random = ~wave, subject='SUBJID', ng = 6, data = df_numeric, B=m1)
# Model with no latent classes (for reference)
```

```{r}
#summary table for model comparison
tab <- as.data.frame(lcmm::summarytable(m1, m2, m3, m4, m5, m6))
```

```{r}
mq1 <- lcmm::hlme(ptg ~ wave + I(wave**2), random = ~wave, subject = "SUBJID", ng = 1, data = df_numeric)
# Models with 2 to 6 latent classes
mq2 <- lcmm::hlme(ptg ~ wave + I(wave**2), mixture = ~wave + I(wave**2), random = ~wave, subject = 'SUBJID', ng = 2, data = df_numeric, B=mq1)
mq3 <- lcmm::hlme(ptg ~ wave + I(wave**2), mixture = ~wave + I(wave**2), random = ~wave, subject = 'SUBJID', ng = 3, data = df_numeric, B=mq1)
mq4 <- lcmm::hlme(ptg ~ wave + I(wave**2), mixture = ~wave + I(wave**2), random = ~wave, subject = 'SUBJID', ng = 4, data = df_numeric, B=mq1)
mq5 <- lcmm::hlme(ptg ~ wave + I(wave**2), mixture = ~wave + I(wave**2), random = ~wave, subject = 'SUBJID', ng = 5, data = df_numeric, B=mq1)
mq6 <- lcmm::hlme(ptg ~ wave + I(wave**2), mixture = ~wave + I(wave**2), random = ~wave, subject = 'SUBJID', ng = 6, data = df_numeric, B=mq1)
```
```{r}
#summary table for model comparison
tab <- as.data.frame(lcmm::summarytable(mq1, mq2, mq3, mq4, mq5, mq6))
```

```{r}
table(m3$pprob[,2])
```
```{r}
table(m2$pprob[,2])
```
```{r}
table(mq3$pprob[,2])
```
```{r}
table(mq2$pprob[,2])
```



#lost toys
```{r}
missing <- df_numeric %>% filter(wave == "wave1") %>% 
  select(category, pts1, ptg1,etv1,Gender,ss1.Race,ss1.Age) 
tab_category_gender <- table(missing$category, missing$Gender)
tab_category_race <- table(missing$category, missing$ss1.Race)
tab_category_age <- table(missing$category, missing$ss1.Age)
chisq.test(tab_category_gender) 
chisq.test(tab_category_race)
t.test(pts1 ~ category, data = missing)
t.test(etv1 ~ category, data = missing)
t.test(ptg1 ~ category, data = missing)
```
```{r}
# Create a contingency table
table_data <- with(data1, table(ss1.Race, category))

# Perform the chi-squared test
test_result <- chisq.test(table_data)

# Print the test result
print(test_result)

```
```{r}
table2 <- table(df_numeric$category, df_numeric$clust2)
chisq.test(table2)
```

```{r}
t.test(wave1_etv ~ category, data = data1)
```


```{r}
ptg_wave1 <- data1 %>% select(contains("ss1.ptg"))
pts_wave1 <- data1 %>% select(contains("ss1.pts"))
alpha(pts_wave1)
```





```{r}
pts_race <- aov(wave1_pts ~ ss1.Race, data = data1)
summary(pts_race)
```
```{r}
ptg_race <- aov(wave1_ptg ~ ss1.Race, data = data1)
summary(ptg_race)
```

```{r}
etv_race <- aov(wave1_etv ~ ss1.Race, data = data1)
summary(etv_race)
```
```{r}
pts_gender <- t.test(wave1_pts ~ Gender, data = data1)
summary(pts_gender)
```
```{r}
ptg_gender <- t.test(wave1_ptg ~ Gender, data = data1)
summary(ptg_gender)
```













```{r}
ggplot(df_long, aes(x = wave, y = ptg, fill = wave)) +
  geom_boxplot() +
  scale_fill_manual(values = c("blue", "red", "yellow"))
```






```{r}
data %>%
  filter(!is.na(ss3.Sex)) %>%
  group_by(ss3.Sex) %>%
  summarise(count = n()) %>%
  mutate(percentage = (count / sum(count)) * 100)
```
```{r}
data %>%
  filter(!is.na(ss4.Sex)) %>%
  group_by(ss4.Sex) %>%
  summarise(count = n()) %>%
  mutate(percentage = (count / sum(count)) * 100)
```
```{r}
data %>%
  filter(!is.na(ss2.Sex)) %>%
  group_by(ss2.Sex) %>%
  summarise(count = n()) %>%
  mutate(percentage = (count / sum(count)) * 100)
```











```{r}
b <- ggplot(df_numeric, aes(x=wave, y=ptg, group=SUBJID, color=SUBJID)) + 
  geom_line() +
  geom_point() 

print(b)
```


```{r}
summary(m2)
```
```{r}
summary(m6)
```


```{r}
pp_m2 <- postprob(m2)
head(pp_m2)
```
```{r}
postprob(m3)
```


## The best fitting model appears to be a quadratic model with a single wave, but that's not necessarily what makes sense with our other intuitions and knowledge of this phenomenon. The second best fitting model is the linear 3 class, closely followed by the quadratic three class.
Model Selection
```{r}
#summary table for model comparison
tab1 <- as.data.frame(lcmm::summarytable(m1, m2, m3, m4, m5, m6))
tab2 <- as.data.frame(lcmm::summarytable(mq1, mq2, mq3, mq4, mq5, mq6))
tab3 <- as.data.frame(lcmm::summarytable(m1, m2, m3, m4, m5, m6, mq1, mq2, mq3, mq4, mq5, mq6))

```
```{r}
summary(m3)
```
```{r}
summary(mq3)
```



```{r}
summary(m6)
```

```{r}
ggplot(tab2, aes(x=G, y=BIC)) +
  geom_point(size = 3, shape=21) + 
  geom_line(size = .3) +
  scale_y_continuous(limits = c(2100,2390)) +
  scale_x_continuous(breaks = c(1:6)) +
  xlab("Number of clusters") +
  theme_classic()
```


```{r}
chisq.posthoc.test(conting_race)
```


```{r}
summary(m2)
summary(m2etv_1)
summary(m2pts_1)
summary(m2pts_full)
summary(m2etv_full)
summary(m2race_full)
```
```{r}
m2gender <- lcmm::hlme(
  fixed = ptg ~ wave + Gender,  # Include ss1.Sex in fixed effects
  mixture = ~ wave + Gender,    # Include ss1.Sex in mixture (latent classes)
  random = ~ wave,                     # Random intercept and slope by wave
  subject = 'SUBJID',                  # Subject ID to specify the clustering of observations
  classmb = ~ Gender,           # Include gun and ss1.Sex for class membership
  ng = 2,                              # Number of latent classes
  data = df_numeric                    # Data frame
)

```
```{r}
m2age <- lcmm::hlme(
  fixed = ptg ~ wave + ss1.Age,  # Include ss1.Sex in fixed effects
  mixture = ~ wave + ss1.Age,    # Include ss1.Sex in mixture (latent classes)
  random = ~ wave,                     # Random intercept and slope by wave
  subject = 'SUBJID',                  # Subject ID to specify the clustering of observations
  classmb = ~ ss1.Age,           # Include gun and ss1.Sex for class membership
  ng = 2,                              # Number of latent classes
  data = df_numeric                    # Data frame
)

```
```{r}
m2race <- lcmm::hlme(
  fixed = ptg ~ wave + ss1.Race,  # Include ss1.Sex in fixed effects
  mixture = ~ wave + ss1.Race,    # Include ss1.Sex in mixture (latent classes)
  random = ~ wave,                     # Random intercept and slope by wave
  subject = 'SUBJID',                  # Subject ID to specify the clustering of observations
  classmb = ~ ss1.Race,           # Include gun and ss1.Sex for class membership
  ng = 2,                              # Number of latent classes
  data = df_numeric                    # Data frame
)

```



Other Models

```{r}
simple_model <- lcmm::hlme(fixed = ptg ~ wave + pts1,
                           mixture = ~pts1,
                           random = ~wave,
                           subject = "SUBJID",
                           ng = 2, 
                           data = df_numeric)

```
```{r}
simple_model1 <- lcmm::hlme(fixed = ptg ~ wave + etv1,
                           mixture = ~etv1,
                           random = ~wave,
                           subject = "SUBJID",
                           ng = 2, 
                           data = df_numeric)

```
```{r}
simple_model2 <- lcmm::hlme(fixed = ptg ~ wave + gun1,
                           mixture = ~gun1,
                           random = ~wave,
                           subject = "SUBJID",
                           ng = 2, 
                           data = df_numeric)

```







```{r}
d <- aov(etv ~ wave, data = df_numeric)
```
```{r}
summary(d)
```

```{r}
e <- aov(gun ~ wave, data = df_numeric)
```
```{r}
summary(e)
```
```{r}
data1 %>% 
  summarise(mean1 = mean(wave1_pts), mean2 = mean(wave2_pts, na.rm = TRUE), mean3 = mean(wave3_pts, na.rm = TRUE))
```

```{r}
f <- aov(pts ~ wave, data = df_numeric)
```
```{r}
summary(f)
```


```{r}
summary(m3)
```
```{r}
summary(m2)
```


```{r}
table(m3$pprob[,2])
```
```{r}
table(m2$pprob[,2])
```
```{r}
table(mq3$pprob[,2])
```
```{r}
table(mq2$pprob[,2])
```
INTRODUCE CLUSTERS BELOW

```{r}
df_numeric <- merge(df_numeric, round(m3$pprob, 3)[, 1:2], by = "SUBJID")
df_numeric$clust3 <- as.factor(df_numeric$class)
df_numeric$class <- NULL
df_numeric <- merge(df_numeric, round(m2$pprob, 3)[, 1:2], by = "SUBJID")
df_numeric$clust2 <- as.factor(df_numeric$class)
df_numeric$class <- NULL
df_numeric <- merge(df_numeric, round(mq3$pprob, 3)[, 1:2], by = "SUBJID")
df_numeric$clustq3 <- as.factor(df_numeric$class)
df_numeric$class <- NULL
df_numeric <- merge(df_numeric, round(mq2$pprob, 3)[, 1:2], by = "SUBJID")
df_numeric$clustq2 <- as.factor(df_numeric$class)
df_numeric$class <- NULL
```

```{r}
summary(m1)
```

```{r}

```

```{r}
conting_gender <- table(df_numeric$Gender, df_numeric$clust2)
conting_race <- table(df_numeric$ss1.Race, df_numeric$clust2)
conting_age <- table(df_numeric$ss1.Age, df_numeric$clust2)
chisq.test(conting_gender)
chisq.test(conting_race)
chisq.test(conting_age)
```

Post-traumatic stress is significantly higher for those in Group 2

```{r}
conting_race
```


```{r}
t.test(pts3 ~ clust2, data = df_numeric)
```

```{r}
t.test(pts1 ~ clust2, data = df_numeric)
```
```{r}
t.test(etv1 ~ clust2, data = df_numeric)
```
Post-traumatic stress at time 3 was also higher for the high increasing post-traumatic growth group

Cluster membership does not significantly differ by gender, age,  or race.

```{r}
ggplot(df_numeric, aes(x=wave, y=ptg, colour = etvclust2, group = SUBJID)) +
  geom_point(size = .1) + geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = etvclust2)) +
  scale_color_manual(values=c( "red","green"))
```
```{r}
ggplot(df_numeric, aes(x=wave, y=ptg, colour = ptsclust2, group = SUBJID)) +
  geom_point(size = .1) + geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = ptsclust2)) +
  scale_color_manual(values=c( "red","green"))
```
```{r}
ggplot(df_numeric, aes(x=wave, y=ptg, colour = clust2, group = SUBJID)) +
  geom_point(size = .1) + geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = clust2)) +
  scale_color_manual(values=c( "red","green"))
```
```{r}
ggplot(df_numeric, aes(x=wave, y=ptg, colour = clust3, group = SUBJID)) +
  geom_point(size = .1) + geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = clust3)) +
  scale_color_manual(values=c( "red","green","blue"))
```

```{r}
df_numeric$clust2
```

Cleaned table
```{r}
library(ggplot2)

ggplot(df_numeric, aes(x=wave, y=ptg, colour = clust2, group = SUBJID)) +
  geom_point(size = .1) +
  geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = clust2)) +
  scale_color_manual(values=c("red","green"), 
                     name = "Class", 
                     breaks = c("1", "2"), 
                     labels = c("Low-Decreasing", "High-Increasing")) +
  labs(
    title = "Two Class Model for Post-Traumatic Growth Trajectories",
    x = "Wave",
    y = "Post-Traumatic Growth"
  ) + 
  theme_minimal()

```
```{r}
ggplot(df_numeric, aes(x=wave, y=ptg, linetype = clustq3, group = SUBJID)) +
  geom_point(size = .1) +
  geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = clust2)) +
  scale_linetype_manual(values=c("solid","dashed"), 
                        name = "Class", 
                        breaks = c("1", "2"), 
                        labels = c("Low-Decreasing", "High-Increasing")) +
  labs(
    title = "Two Class Model for Post-Traumatic Growth Trajectories",
    x = "Wave",
    y = "Post-Traumatic Growth"
  ) + 
  theme_minimal()

ggplot(df_numeric, aes(x=wave, y=ptg, linetype = clustq3, group = SUBJID)) +
  geom_point(linewidth = .1) + 
  geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = clust3)) +
  scale_color_manual(values=c( "red","green","blue"))

ggplot(df_numeric, aes(x=wave, y=ptg, linetype = clust3, group = SUBJID)) +
  geom_point(size = .1) + 
  geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = clustq3)) +
  scale_color_manual(values=c( "red","green","blue"))
```
```{r}
ggplot(df_numeric, aes(x=wave, y=ptg, linetype = clust2, group = SUBJID)) +
  geom_point(size = .1) +
  geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = clust2)) +
  scale_linetype_manual(values=c("solid","dashed"), 
                        name = "Class", 
                        breaks = c("1", "2"), 
                        labels = c("Low-Decreasing", "High-Increasing")) +
  labs(
    title = "Two Class Model for Post-Traumatic Growth Trajectories",
    x = "Wave",
    y = "Post-Traumatic Growth"
  ) + 
  theme_minimal()
```


```{r}
ggplot(df_numeric, aes(x=wave, y=ptg, colour = etvclust2, group = SUBJID)) +
  geom_point(size = .1) +
  geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = etvclust2)) +
  scale_color_manual(values=c("red","green"), 
                     name = "Class", 
                     breaks = c("1", "2"), 
                     labels = c("Low-Decreasing", "High-Increasing")) +
  labs(
    title = "Two Class Model for Post-Traumatic Growth Trajectories with ETV Covariate",
    x = "Wave",
    y = "Post-Traumatic Growth"
  ) + 
  theme_minimal()

```

```{r}
library(ggplot2)

ggplot(df_numeric, aes(x=wave, y=ptg, colour = ptsclust2, group = SUBJID)) +
  geom_point(size = .1) +
  geom_line(size = .1) +
  geom_smooth(se = FALSE, method = "loess", aes(group = ptsclust2)) +
  scale_color_manual(values=c("red","green"), 
                     name = "Class", 
                     breaks = c("1", "2"), 
                     labels = c("Low-Decreasing", "High-Increasing")) +
  labs(
    title = "Two Class Model for Post-Traumatic Growth Trajectories with PTS Covariate",
    x = "Wave",
    y = "Post-Traumatic Growth"
  ) + 
  theme_minimal()

```

```{r}
summary(m2)
```


Age and Gender Differences

```{r}
t.test(ptg ~ Gender, data = df_numeric)
```
```{r}
unique(df_numeric$Gender)
```


Models
```{r}
#intercept only model
baseline_w1 <- lmer(ptg ~ 1 + (1 | SUBJID), data = df_merged, REML = FALSE)
```
```{r}
fixed_pts_wave1 <- lmer(ptg ~ 1 + pts + (1 | SUBJID), data = df_merged, REML = FALSE)
```
```{r}
#random fixed pts effect
ranfixed_pts_wave1 <- lmer(ptg ~ 1 + pts + (1 + pts | SUBJID), data = df_merged, REML = FALSE)
#getting result boundary (singular) fit: see ?isSingular
```
```{r}
#fixed etv effect
fixed_etv_wave1 <- lmer(ptg ~ 1 + etv + (1 | SUBJID), data = df_merged, REML = FALSE)
```
```{r}
#random fixed etv effect
ranfixed_etv_wave1 <- lmer(ptg ~ 1 + etv + (1 + etv | SUBJID), data = df_merged, REML = FALSE)
```

```{r}
etv_mod_c <- anova(baseline_w1, ranfixed_etv_wave1, fixed_etv_wave1)
```
```{r}
print(etv_mod_c)
```
```{r}
pts_mod_c <- anova(baseline_w1, ranfixed_pts_wave1, fixed_pts_wave1)
```
```{r}
print(pts_mod_c)
```


```{r}
summary(baseline_w1)
```

```{r}
summary(fixed_pts_wave1)
```
With PTS at 0, the average PTG is 2.55. Every time PTS goes up by a point, the average PTG score rises by .38. The standard deviation for the residual dropped a very small amount.

```{r}
summary(ranfixed_pts_wave1)
```
```{r}
summary(fixed_etv_wave1)
```

```{r}
summary(ranfixed_etv_wave1)
```
```{r}
#intercept only model
baseline <- lmer(ptg ~ 1 + (1 | SUBJID), data = df_clean)
```
```{r}
#fixed pts effect
fixed_pts <- lmer(ptg ~ 1 + pts + (1 | SUBJID), data = df_clean)
```
```{r}
#random fixed pts effect
ranfixed_pts <- lmer(ptg ~ 1 + pts + (1 + pts | SUBJID), data = df_clean)
```

```{r}
#fixed etv effect
fixed_etv <- lmer(ptg ~ 1 + etv + (1 | SUBJID), data = df_clean)
```
```{r}
#random fixed etv effect
ranfixed_etv <- lmer(ptg ~ 1 + etv + (1 + etv | SUBJID), data = df_clean)
```

OKAY! The random and fixed effects model for PTS and ETV offers the best fit. 

```{r}
summary(baseline)
```
The mean PTG is 3.44.
```{r}
summary(fixed_pts)
```
With PTS at 0, the average PTG is 2.55. Every time PTS goes up by a point, the average PTG score rises by .38. The standard deviation for the residual dropped a very small amount.

```{r}
summary(ranfixed_pts)
```
```{r}
summary(fixed_etv)
```

```{r}
summary(ranfixed_etv)
```
For ETV, a fixed effect model appears worse than baseline, but a linear mixed effect model with random effects improves on the model.

```{r}
summary(baseline)
```
The mean PTG is 3.44.




```{r}
data3 <- na.omit(data.frame(data1$wave1_ptg, data1$wave2_ptg, data1$wave3_ptg))
```
```{r}
ptg.cov <- cov(data3)
```
```{r}
data3
```

```{r}
pc1_ptg <- lm(wave2_ptg ~ wave1_ptg + wave1_pts, data = data1)
pc2_ptg <- lm(wave3_ptg ~ wave2_ptg + wave2_pts, data = data1)
pc1_pts <- lm(wave2_pts ~ wave1_ptg + wave1_pts, data = data1)
pc2_pts <- lm(wave3_pts ~ wave2_ptg + wave2_pts, data = data1)
```
```{r}
summary(pc1_pts)
```




3. Latent class mixture model



```{r}
ggplot(data = df_numeric, aes(x = pts, y = ptg, color = factor(wave))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "PTS", y = "PTG", color = "Wave") +
  theme_minimal()
```
```{r}
ggplot(data = df_numeric, aes(x = etv, y = ptg, color = factor(wave))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "ETV", y = "PTG", color = "Wave") +
  theme_minimal()
```

```{r}
df_numeric <- merge(df_numeric, round(m2$pprob, 3)[, 1:2], by = "SUBJID")
df_numeric$clust2 <- as.factor(df_numeric$class)
df_numeric$class <- NULL
df_numeric <- merge(df_numeric, round(m2etv_1$pprob, 3)[, 1:2], by = "SUBJID")
df_numeric$etvclust2 <- as.factor(df_numeric$class)
df_numeric$class <- NULL
df_numeric <- merge(df_numeric, round(m2pts_1$pprob, 3)[, 1:2], by = "SUBJID")
df_numeric$ptsclust2 <- as.factor(df_numeric$class)
df_numeric$class <- NULL
df_numeric$wave <- as.numeric(sub("wave", "", df_numeric$wave))

```









